---
title: "Human-Animal Contact Networks: Full Pipeline"
author: "Lev Kolinski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
library(tidyverse)
library(here)
library(igraph)
library(statnet)
library(ergm)
library(ergMargins)
library(metafor)
library(ggpubr)
library(ineq)
library(kableExtra)
library(broom)

set.seed(1234)
```

---

# 1. Load Raw Data

```{r load-data}
demographic_health_data <- read.csv(
  "/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Demographic_Health.csv"
)
animal_interaction_data <- read.csv(
  "/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Animal_Interaction.csv"
)
```

---

# 2. Clean and Merge Data

```{r clean-merge}
demographic_health_data_cleaned <- demographic_health_data |>
  mutate(across(
    starts_with(c("own_", "banking", "grew_crops", "crops")),
    ~ ifelse(. == "Yes", 1L, 0L)
  )) |>
  select(
    survey_date, social_netid, village, age, gender, school_level,
    household_size, main_activity,
    starts_with("material_"), starts_with("own_"),
    crops_vanilla, landsize_in_daba
  )

animal_interaction_data_cleaned <- animal_interaction_data |>
  mutate(across(-starts_with("social_netid"), ~ ifelse(. == "Yes", 1L, 0L))) |>
  select(-starts_with("symptoms"), -starts_with("action_"),
         -contains("frequency"))

merged_df_raw <- inner_join(
  demographic_health_data_cleaned,
  animal_interaction_data_cleaned,
  by = "social_netid"
)

cat("N after inner_join:", nrow(merged_df_raw), "\n")
cat("Unique IDs:", length(unique(merged_df_raw$social_netid)), "\n")

```

---

# 3. Feature Engineering

## 3a. Helper Functions

```{r feature-functions}

safe_scale <- function(x, label = "") {
  s <- sd(x, na.rm = TRUE)
  if (is.na(s) || s == 0) {
    if (nchar(label) > 0)
      message("safe_scale: zero/NA variance in '", label,
              "' — returning 0 for non-NA rows")
    return(ifelse(is.na(x), NA_real_, 0))  # was rep(0, length(x))
  }
  as.numeric(scale(x))
}

# ------------------------------------------------------------------
# prepare_village_data(): clean, index, and scale one village subset.
#
# village_filter: character vector of village names to retain, or
#                 NULL (default) to keep all villages.
#
# Scaling is done WITHIN the filtered subset (within-village SDs).
# safe_scale() prevents zero-variance NaN propagation.
# ------------------------------------------------------------------
prepare_village_data <- function(df, village_filter = NULL) {

  if (!is.null(village_filter)) {
    df <- df |> filter(village %in% village_filter)
  }

  vname <- paste(village_filter, collapse = "/")

  # --- ordinal housing material indices (unscaled) ---
  # NOTE: "ravenala" — not "ravinala" — is the correct spelling.
  # A misspelling here causes case_when to hit TRUE ~ NA_real_ for all
  # ravenala rows, producing an all-NA floor index, which propagates
  # NaN through house_sol and causes mass vertex deletion downstream.
  df <- df |>
    mutate(
      material_wall_index = case_when(
        material_wall %in% c("bamboo", "rafia", "ravenala",
                             "mud", "compacted_earth") ~ 0,
        material_wall == "wood_planks"                 ~ 1,
        material_wall %in% c("brick_unfired",
                             "brick_fired")            ~ 2,
        material_wall == "metal_sheets"                ~ 3,
        material_wall == "cement"                      ~ 4,
        TRUE                                           ~ NA_real_
      ),
      material_roof_index = case_when(
        material_roof %in% c("bamboo", "thatch") ~ 0,
        material_roof == "wood_planks"          ~ 0.5,
        material_roof == "metal_sheets"          ~ 1,
        material_roof == "cement"                ~ 2,
        TRUE                                     ~ NA_real_
      ),
      material_floor_index = case_when(
        material_floor %in% c("dirt", "bamboo",
                              "rafia", "ravenala") ~ 0,
        material_floor == "wood_planks"            ~ 1,
        material_floor == "cement"                 ~ 2,
        TRUE                                       ~ NA_real_
      )
    )

  # --- scale everything within this village subset ---
  df |>
    mutate(
      house_sol = safe_scale(
        material_wall_index + material_roof_index + material_floor_index,
        paste(vname, "house_sol")
      ),
      commercial_goods = safe_scale(
        own_cellphone + own_tv + own_bicycle + own_refrigerator +
          own_motorcycle + own_computer + own_generator,
        paste(vname, "commercial_goods")
      ),
      # FIX: rowSums inside mutate gives per-row totals
      total_animal_interactions = rowSums(
        across(pet_dogs:dead_goats_sheep), na.rm = TRUE
      ),
      employment_category = case_when(
        grepl("farm_crops|farm_mixed", main_activity)      ~ "Farmer",
        grepl("student|teacher", main_activity)            ~ "Education",
        grepl("unemployed|retired_teacher", main_activity) ~ "Unemployed",
        TRUE                                               ~ "Other"
      ),
      # FIX: rowSums gives per-person high-risk exposure counts
      num_high_risk_exposures = rowSums(
        across(contains(c("shared_water", "scratched_bitten", "feces",
                          "dead", "raw_undercooked", "eaten_sick"))),
        na.rm = TRUE
      ),
      high_risk_exposures_binary = ifelse(
        num_high_risk_exposures == 0L, 0L, 1L
      ),
      school_level_numbered = safe_scale(
        case_when(
          school_level == "None"      ~ 0L,
          school_level == "Primary"   ~ 1L,
          school_level == "Secondary" ~ 2L,
          school_level == "Higher"    ~ 3L,
          TRUE                        ~ NA_integer_
        ),
        paste(vname, "school_level")
      ),
      age              = safe_scale(age,
                                    paste(vname, "age")),
      landsize_in_daba = safe_scale(landsize_in_daba,
                                    paste(vname, "land_size")),
      household_size   = safe_scale(household_size,
                                    paste(vname, "household_size"))
    ) |>
    rename(grew_vanilla = crops_vanilla)
}
```

## 3b. Apply Feature Engineering (Within-Village Scaling)

```{r apply-features}
merged_df <- bind_rows(
  prepare_village_data(merged_df_raw,
                       village_filter = c("Ampandrana", "Andatsakala")),
  prepare_village_data(merged_df_raw,
                       village_filter = "Mandena"),
  prepare_village_data(merged_df_raw,
                       village_filter = "Sarahandrano")
) |>
  mutate(survey_date = as.Date(survey_date))

cat("Total respondents:", length(unique(merged_df$social_netid)), "\n")
```

# 4. Build Bipartite Networks

## 4a. Animal Lists and Categories

```{r animal-lists}
animals_full      <- c("rodents", "poultry", "dogs", "cats",
                        "domestic_pigs", "goats_sheep", "cows",
                        "wild_birds", "bush_pigs", "lemurs",
                        "tenrecs", "carnivores")
animals_high_risk <- animals_full

# Full network: 3 animal categories
domesticated_animals <- c("poultry", "dogs", "domestic_pigs",
                           "cows", "goats_sheep", "cats")
rodent_animals       <- "rodents"
wild_animals         <- c("bush_pigs", "carnivores", "tenrecs",
                           "wild_birds", "lemurs")

# High-risk network: domesticated split into two groups
domesticated_animals_1 <- c("poultry", "domestic_pigs", "cows")
domesticated_animals_2 <- c("dogs", "goats_sheep", "cats")

village_groups <- list(
  ampandrana_andatsakala = c("Ampandrana", "Andatsakala"),
  mandena                = "Mandena",
  sarahandrano           = "Sarahandrano"
)
```

## 4b. Helper: Build Edge List

```{r edge-list-function}

build_edge_list <- function(df, animals, col_pattern_fn) {
  edge_dfs <- lapply(animals, function(animal) {
    cols <- col_pattern_fn(animal, df)
    if (length(cols) == 0) return(NULL)
    contact_mat <- as.data.frame(df[cols])
    ids <- df$social_netid[rowSums(contact_mat, na.rm = TRUE) > 0]
    if (length(ids) == 0) return(NULL)
    data.frame(social_netid = ids, animal = animal,
               stringsAsFactors = FALSE)
  })

  result <- bind_rows(edge_dfs)
  if (nrow(result) == 0 || !"social_netid" %in% names(result)) {
    return(data.frame(social_netid = character(0),
                      animal       = character(0),
                      stringsAsFactors = FALSE))
  }
  arrange(result, social_netid)
}

# Full network: any contact column for this animal
full_col_pattern <- function(animal, df) {
  grep(paste0("_", animal, "$|_", animal, "_"), names(df), value = TRUE)
}

# High-risk network: exposure type precedes animal name
# e.g. "feces_rodents", "scratched_bitten_dogs", "slaughtered_poultry"
high_risk_col_pattern <- function(animal, df) {
  grep(
    paste0("(feces|shared_water|scratched_bitten|cooked_handled|",
           "raw_undercooked|eaten_sick|slaughtered)_", animal),
    names(df), value = TRUE
  )
}
```

## 4c. Helper: Build igraph Bipartite Object

```{r build-graph-function}
build_bipartite_graph <- function(edge_list, all_humans, all_animals, df) {

  g <- graph_from_data_frame(edge_list, directed = FALSE)

  # Add back isolates for both modes
  human_isolates  <- setdiff(all_humans,  edge_list$social_netid)
  animal_isolates <- setdiff(all_animals, edge_list$animal)
  g <- add_vertices(g, nv = length(human_isolates),  name = human_isolates)
  g <- add_vertices(g, nv = length(animal_isolates), name = animal_isolates)

  # Set vertex types: TRUE = human, FALSE = animal
  animal_vids <- which(V(g)$name %in% all_animals)
  V(g)$type   <- TRUE
  V(g)$type[animal_vids] <- FALSE
  human_vids  <- which(V(g)$type)

  # Attach human attributes
  attach_attr <- function(g, attr_name, col, vids) {
    vals <- rep(NA, vcount(g))
    vals[vids] <- col[match(V(g)$name[vids], df$social_netid)]
    vertex_attr(g, attr_name) <- vals
    g
  }

  g <- attach_attr(g, "village",          df$village,               human_vids)
  g <- attach_attr(g, "age",              df$age,                   human_vids)
  g <- attach_attr(g, "gender",           df$gender,                human_vids)
  g <- attach_attr(g, "commercial_goods", df$commercial_goods,      human_vids)
  g <- attach_attr(g, "house_sol",        df$house_sol,             human_vids)
  g <- attach_attr(g, "grew_vanilla",     df$grew_vanilla,          human_vids)
  g <- attach_attr(g, "land_size",        df$landsize_in_daba,      human_vids)
  g <- attach_attr(g, "household_size",   df$household_size,        human_vids)
  g <- attach_attr(g, "school_level",     df$school_level_numbered, human_vids)

  V(g)$animal_name <- "Human"
  V(g)$animal_name[animal_vids] <- V(g)$name[animal_vids]

  g
}
```

## 4d. Helper: Village Subgraph

```{r village-subgraph-function}
# Animal vertices (NA village) are always included regardless of
# whether they have edges within that village — zero-edge animal
# nodes are retained as isolates so all 12 types are present
# in every village network.
make_village_subgraph <- function(g, village_names) {
  vids  <- which(V(g)$village %in% village_names | is.na(V(g)$village))
  sub_g <- subgraph(g, vids)
  V(sub_g)$degree <- igraph::degree(sub_g)
  sub_g
}
```

## 4e. Helper: Assign Animal Categories

```{r animal-category-function}
assign_animal_categories <- function(g, category_map) {
  cats <- rep(NA_character_, vcount(g))
  for (cat_label in names(category_map)) {
    cats[V(g)$name %in% category_map[[cat_label]]] <- cat_label
  }
  V(g)$animal_category <- cats
  g
}

cat_map_full <- list(
  domesticated = domesticated_animals,
  rodent       = rodent_animals,
  wild         = wild_animals
)

cat_map_high_risk <- list(
  domesticated_1 = domesticated_animals_1,
  domesticated_2 = domesticated_animals_2,
  rodent         = rodent_animals,
  wild           = wild_animals
)
```

## 4f. Build Full Network

```{r build-full-network}
edge_list_full   <- build_edge_list(merged_df, animals_full,
                                    full_col_pattern)
bipartite_graph_full <- build_bipartite_graph(
  edge_list   = edge_list_full,
  all_humans  = unique(merged_df$social_netid),
  all_animals = animals_full,
  df          = merged_df
)

bg_full_amp <- assign_animal_categories(
  make_village_subgraph(bipartite_graph_full,
                        village_groups$ampandrana_andatsakala),
  cat_map_full)
bg_full_man <- assign_animal_categories(
  make_village_subgraph(bipartite_graph_full, village_groups$mandena),
  cat_map_full)
bg_full_sar <- assign_animal_categories(
  make_village_subgraph(bipartite_graph_full, village_groups$sarahandrano),
  cat_map_full)

cat("Full network — vertices:", vcount(bipartite_graph_full),
    "| edges:", ecount(bipartite_graph_full), "\n")
```

## 4g. Build High-Risk Network

```{r build-high-risk-network}
for (animal in animals_high_risk) {
  cols <- high_risk_col_pattern(animal, merged_df)
  if (length(cols) == 0) {
    warning("No high-risk columns found for: ", animal)
    merged_df[[paste0(animal, "_edge_high_risk")]] <- 0L
  } else {
    merged_df[[paste0(animal, "_edge_high_risk")]] <-
      as.integer(rowSums(merged_df[cols], na.rm = TRUE) > 0)
  }
}

edge_list_high_risk <- build_edge_list(
  merged_df, animals_high_risk,
  function(animal, df) paste0(animal, "_edge_high_risk")
)
bipartite_graph_high_risk <- build_bipartite_graph(
  edge_list   = edge_list_high_risk,
  all_humans  = unique(merged_df$social_netid),
  all_animals = animals_high_risk,
  df          = merged_df
)

bg_hr_amp <- assign_animal_categories(
  make_village_subgraph(bipartite_graph_high_risk,
                        village_groups$ampandrana_andatsakala),
  cat_map_high_risk)
bg_hr_man <- assign_animal_categories(
  make_village_subgraph(bipartite_graph_high_risk,
                        village_groups$mandena),
  cat_map_high_risk)
bg_hr_sar <- assign_animal_categories(
  make_village_subgraph(bipartite_graph_high_risk,
                        village_groups$sarahandrano),
  cat_map_high_risk)

cat("High-risk network — vertices:", vcount(bipartite_graph_high_risk),
    "| edges:", ecount(bipartite_graph_high_risk), "\n")
```

---

# 5. Convert igraph to statnet and Run ERGMs

## 5a. Conversion Helper

```{r conversion-function}
# ------------------------------------------------------------------
# Converts a bipartite igraph to a statnet network object,
# re-attaches node attributes, and removes human vertices with
# missing covariate data.
# ------------------------------------------------------------------
igraph_to_statnet <- function(ig) {

  node_df <- data.frame(
    id               = V(ig)$name,
    type             = V(ig)$type,
    village          = V(ig)$village,
    age              = V(ig)$age,
    gender           = V(ig)$gender,
    commercial_goods = V(ig)$commercial_goods,
    house_sol        = V(ig)$house_sol,
    grew_vanilla     = V(ig)$grew_vanilla,
    land_size        = V(ig)$land_size,
    household_size   = V(ig)$household_size,
    school_level     = V(ig)$school_level,
    degree           = V(ig)$degree,
    animal_name      = V(ig)$animal_name,
    animal_category  = V(ig)$animal_category,
    stringsAsFactors = FALSE
  )

  human_ids  <- node_df$id[ node_df$type]
  animal_ids <- node_df$id[!node_df$type]
  n_humans   <- length(human_ids)
  n_animals  <- length(animal_ids)

  
  biadj <- matrix(
    0L,
    nrow     = n_animals,
    ncol     = n_humans,
    dimnames = list(animal_ids, human_ids)
  )
  
  el <- igraph::as_edgelist(ig)
  for (i in seq_len(nrow(el))) {
    u <- el[i, 1]; v <- el[i, 2]
    if      (u %in% animal_ids && v %in% human_ids) biadj[u, v] <- 1L
    else if (v %in% animal_ids && u %in% human_ids) biadj[v, u] <- 1L
  }
  
  stopifnot(
    "biadj rows are not all animal IDs" =
      all(rownames(biadj) %in% animal_ids),
    "biadj cols are not all human IDs" =
      all(colnames(biadj) %in% human_ids)
  )
  
  sn <- network(biadj, bipartite = n_animals, directed = FALSE)

  stopifnot(
    "statnet vertex count does not match igraph" =
      network.size(sn) == n_humans + n_animals,
    "statnet edge count does not match igraph"   =
      network.edgecount(sn) == igraph::ecount(ig)
  )

  cat("Conversion OK:", network.size(sn), "vertices |",
      network.edgecount(sn), "edges\n")

  statnet_ids     <- as.character(sn %v% "vertex.names")
  node_df_ordered <- node_df[match(statnet_ids, node_df$id), ]

  if (!isTRUE(all(node_df_ordered$id == statnet_ids))) {
    mismatch_idx <- which(node_df_ordered$id != statnet_ids |
                            is.na(node_df_ordered$id))
    stop("Vertex ID mismatch after reordering!\n",
         "Positions: ",    paste(mismatch_idx, collapse = ", "), "\n",
         "statnet wants: ", paste(statnet_ids[mismatch_idx],      collapse = ", "), "\n",
         "node_df has:   ", paste(node_df_ordered$id[mismatch_idx], collapse = ", "))
  }

  sn %v% "village"          <- node_df_ordered$village
  sn %v% "age"              <- node_df_ordered$age
  sn %v% "gender"           <- node_df_ordered$gender
  sn %v% "commercial_goods" <- node_df_ordered$commercial_goods
  sn %v% "house_sol"        <- node_df_ordered$house_sol
  sn %v% "grew_vanilla"     <- node_df_ordered$grew_vanilla
  sn %v% "land_size"        <- node_df_ordered$land_size
  sn %v% "household_size"   <- node_df_ordered$household_size
  sn %v% "school_level"     <- node_df_ordered$school_level
  sn %v% "degree"           <- node_df_ordered$degree
  sn %v% "animal_name"      <- node_df_ordered$animal_name
  sn %v% "animal_category"  <- node_df_ordered$animal_category

  # Drop humans with missing covariates only — never animal vertices
  covariate_attrs <- c("age", "gender", "commercial_goods", "house_sol",
                       "grew_vanilla", "land_size", "household_size",
                       "school_level")
  animal_vids <- which(sn %v% "animal_name" != "Human")
  vertices_with_na <- Reduce(union, lapply(covariate_attrs, function(a)
    which(is.na(sn %v% a))))
  vertices_to_drop <- setdiff(vertices_with_na, animal_vids)

  if (length(vertices_to_drop) > 0) {
    cat("Dropped vertices:\n")
    print((sn %v% "vertex.names")[vertices_to_drop])

    cat("Dropping", length(vertices_to_drop),
       "human vertices with missing covariates.\n")
    sn <- delete.vertices(sn, vertices_to_drop)
  }

  sn
}
```

## 5b. Convert All Networks

```{r convert-networks, message=TRUE, warning=FALSE}
sn_full_amp <- igraph_to_statnet(bg_full_amp)
sn_full_man <- igraph_to_statnet(bg_full_man)
sn_full_sar <- igraph_to_statnet(bg_full_sar)

sn_hr_amp   <- igraph_to_statnet(bg_hr_amp)
sn_hr_man   <- igraph_to_statnet(bg_hr_man)
sn_hr_sar   <- igraph_to_statnet(bg_hr_sar)
```

## 5c. Run ERGMs — Full Network

```{r ergm-full, cache=TRUE}
ergm_full_amp <- ergm(
  sn_full_amp ~
    edges + b2cov("age") + b2factor("gender") +
    b2cov("commercial_goods") + b2cov("house_sol") +
    b2factor("grew_vanilla") + b2cov("land_size") +
    b2cov("household_size") + b2cov("school_level") +
    b1factor("animal_category") +
    b1factor("animal_category"):b2cov("commercial_goods") +
    b1factor("animal_category"):b2cov("house_sol") +
    b1factor("animal_category"):b2factor("grew_vanilla") +
    b2degree(2) + b2degree(3) + b2degree(4) +
    b2degree(5) + b2degree(6),
  control = control.ergm(seed = 1234)
)

#save(ergm_full_amp, file = "/Users/levkolinski/Desktop/human_animal_networks/self_reported_networks/Results/ergm_full_village_a.RData")


ergm_full_man <- ergm(
  sn_full_man ~
    edges + b2cov("age") + b2factor("gender") +
    b2cov("commercial_goods") + b2cov("house_sol") +
    b2factor("grew_vanilla") + b2cov("land_size") +
    b2cov("household_size") + b2cov("school_level") +
    b1factor("animal_category") +
    b1factor("animal_category"):b2cov("commercial_goods") +
    b1factor("animal_category"):b2cov("house_sol") +
    b1factor("animal_category"):b2factor("grew_vanilla") +
    b2degree(2) + b2degree(3) + b2degree(4) +
    b2degree(5) + b2degree(6) + b2degree(7) + b2degree(8),
  control = control.ergm(seed = 123)
)

#save(ergm_full_man, file = "/Users/levkolinski/Desktop/human_animal_networks/self_reported_networks/Results/ergm_full_village_m.RData")


ergm_full_sar <- ergm(
  sn_full_sar ~
    edges + b2cov("age") + b2factor("gender") +
    b2cov("commercial_goods") + b2cov("house_sol") +
    b2factor("grew_vanilla") + b2cov("land_size") +
    b2cov("household_size") + b2cov("school_level") +
    b1factor("animal_category") +
    b1factor("animal_category"):b2cov("commercial_goods") +
    b1factor("animal_category"):b2cov("house_sol") +
    b1factor("animal_category"):b2factor("grew_vanilla") +
    b2degree(3) + b2degree(4) + b2degree(5) +
    b2degree(6) + b2degree(7) + b2degree(8),
  control = control.ergm(seed = 123)
)

#save(ergm_full_sar, file = "/Users/levkolinski/Desktop/human_animal_networks/self_reported_networks/Results/ergm_full_village_s.RData")

```

## 5d. Run ERGMs — High-Risk Network

```{r ergm-high-risk, cache=TRUE}
ergm_hr_amp <- ergm(
  sn_hr_amp ~
    edges + b2cov("age") + b2factor("gender") +
    b2cov("commercial_goods") + b2cov("house_sol") +
    b2factor("grew_vanilla") + b2cov("land_size") +
    b2cov("household_size") + b2cov("school_level") +
    b1factor("animal_category") +
    b1factor("animal_category"):b2cov("commercial_goods") +
    b1factor("animal_category"):b2cov("house_sol") +
    b1factor("animal_category"):b2factor("grew_vanilla") +
    b2degree(1) + b2degree(2) + b2degree(4) +
    b2degree(5) + b2degree(6) + b2degree(8),
  control = control.ergm(seed = 1234)
)

#save(ergm_hr_amp, file = "/Users/levkolinski/Desktop/human_animal_networks/self_reported_networks/Results/ergm_hr_village_a.RData")


ergm_hr_man <- ergm(
  sn_hr_man ~
    edges + b2cov("age") + b2factor("gender") +
    b2cov("commercial_goods") + b2cov("house_sol") +
    b2factor("grew_vanilla") + b2cov("land_size") +
    b2cov("household_size") + b2cov("school_level") +
    b1factor("animal_category") +
    b1factor("animal_category"):b2cov("commercial_goods") +
    b1factor("animal_category"):b2cov("house_sol") +
    b1factor("animal_category"):b2factor("grew_vanilla") +
    b2degree(2) + b2degree(3) + b2degree(4) +
    b2degree(5) + b2degree(6) + b2degree(7),
  control = control.ergm(seed = 123)
)

#save(ergm_hr_man, file = "/Users/levkolinski/Desktop/human_animal_networks/self_reported_networks/Results/ergm_hr_village_m.RData")


ergm_hr_sar <- ergm(
  sn_hr_sar ~
    edges + b2cov("age") + b2factor("gender") +
    b2cov("commercial_goods") + b2cov("house_sol") +
    b2factor("grew_vanilla") + b2cov("land_size") +
    b2cov("household_size") + b2cov("school_level") +
    b1factor("animal_category") +
    b1factor("animal_category"):b2cov("commercial_goods") +
    b1factor("animal_category"):b2cov("house_sol") +
    b1factor("animal_category"):b2factor("grew_vanilla") +
    b2degree(2) + b2degree(3) + b2degree(4) +
    b2degree(5) + b2degree(6) + b2degree(7) + b2degree(8),
  control = control.ergm(seed = 1234)
)

#save(ergm_hr_sar, file = "/Users/levkolinski/Desktop/human_animal_networks/self_reported_networks/Results/ergm_hr_village_s.RData")
```

## 5e. MCMC Diagnostics and GOF

```{r diagnostics}
vif.ergm(ergm_full_amp); mcmc.diagnostics(ergm_full_amp); ergm:::plot.gof(gof(ergm_full_amp))
vif.ergm(ergm_full_man); mcmc.diagnostics(ergm_full_man); ergm:::plot.gof(gof(ergm_full_man))
vif.ergm(ergm_full_sar); mcmc.diagnostics(ergm_full_sar); ergm:::plot.gof(gof(ergm_full_sar))

vif.ergm(ergm_hr_amp); mcmc.diagnostics(ergm_hr_amp); ergm:::plot.gof(gof(ergm_hr_amp))
vif.ergm(ergm_hr_man); mcmc.diagnostics(ergm_hr_man); ergm:::plot.gof(gof(ergm_hr_man))
vif.ergm(ergm_hr_sar); mcmc.diagnostics(ergm_hr_sar); ergm:::plot.gof(gof(ergm_hr_sar))
```

---

# 6. Pool Effect Sizes via Meta-Analysis

## 6a. Helpers

```{r meta-helpers}

build_summary_df <- function(ergm_list, term_labels) {
  bind_rows(lapply(names(ergm_list), function(vname) {
    coef_df <- as.data.frame(
      summary(ergm_list[[vname]])$coefficients
    ) |>
      rownames_to_column("raw_term") |>
      filter(!grepl("degree", raw_term))

    if (nrow(coef_df) != length(term_labels))
      stop("Term count mismatch for village: ", vname,
           " (", nrow(coef_df), " vs ", length(term_labels), ")")

    coef_df$Variable <- term_labels
    coef_df$Village  <- vname
    coef_df
  })) |>
    relocate(Variable, Village, .before = Estimate)
}

# ------------------------------------------------------------------
# pool_one() / pool_all(): fixed-effects meta-analysis.
# NOTE: k=3 sites — FE assumes true homogeneity. Consider REML to
# acknowledge between-village variance.
# ------------------------------------------------------------------
pool_one <- function(summary_df, variable) {
  d   <- summary_df |> filter(Variable == variable)
  fit <- rma(yi = d$Estimate, sei = d$`Std. Error`, method = "FE")
  data.frame(Variable = variable, Estimate = as.numeric(fit$beta),
             Lower = fit$ci.lb, Upper = fit$ci.ub,
             `Pr(>|z|)` = fit$pval, SE = fit$se,
             N = nrow(d), check.names = FALSE)
}

pool_all <- function(summary_df) {
  vars <- setdiff(unique(summary_df$Variable), "Edges")
  bind_rows(lapply(vars, \(v) pool_one(summary_df, v))) |>
    mutate(lower_ci    = Estimate - 1.96        * SE,
           upper_ci    = Estimate + 1.96        * SE,
           lower_ci_90 = Estimate - qnorm(0.95) * SE,
           upper_ci_90 = Estimate + qnorm(0.95) * SE)
}

# ------------------------------------------------------------------
# Plotting helpers
# ------------------------------------------------------------------
coef_plot <- function(df, var_levels, color_var = NULL) {
  df$Variable <- factor(df$Variable, levels = rev(var_levels))
  aes_base <- if (!is.null(color_var))
    aes(x = Estimate, y = Variable, color = .data[[color_var]])
  else
    aes(x = Estimate, y = Variable)

  ggplot(df, aes_base) +
    geom_point(position = position_dodge(0.5), size = 3) +
    geom_linerange(aes(xmin = lower_ci,    xmax = upper_ci),
                   position = position_dodge(0.5), linewidth = 0.4) +
    geom_linerange(aes(xmin = lower_ci_90, xmax = upper_ci_90),
                   position = position_dodge(0.5), linewidth = 1.2) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
    labs(x = "Estimated Effect", y = NULL) +
    theme_classic() +
    theme(legend.position = "top", axis.title.x = element_text(size = 12))
}

rename_vars <- function(df) {
  df |> mutate(Variable = str_replace_all(
    Variable, c("CGSOL" = "Goods", "HSOL" = "House Mat.",
                "Gender\\[Male\\]" = "Gender[Man]",
                "Rodents" = "Rodent")))
}

rename_villages <- function(df) {
  df |> mutate(Village = recode(Village,
                                "Ampandrana"   = "A",
                                "Mandena"      = "M",
                                "Sarahandrano" = "S"))
}

add_ci_cols <- function(df, se_col = "SE") {
  df |> mutate(
    lower_ci    = Estimate - 1.96        * .data[[se_col]],
    upper_ci    = Estimate + 1.96        * .data[[se_col]],
    lower_ci_90 = Estimate - qnorm(0.95) * .data[[se_col]],
    upper_ci_90 = Estimate + qnorm(0.95) * .data[[se_col]]
  )
}

people_levels <- c("House Mat.", "Goods", "Vanilla", "Land Size",
                   "Household Size", "School Level", "Gender[Man]", "Age")
```

## 6b. Full Network: Pool and Plot

```{r full-network-pool-plot}
full_term_labels <- c(
  "Edges", "Age", "Gender[Male]", "CGSOL", "HSOL", "Vanilla",
  "Land Size", "Household Size", "School Level",
  "Rodents", "Wild",
  "CGSOL:Rodents", "CGSOL:Wild", "HSOL:Rodents", "HSOL:Wild",
  "Vanilla:Rodents", "Vanilla:Wild"
)

summary_df_full <- build_summary_df(
  list(Ampandrana = ergm_full_amp, Mandena = ergm_full_man,
       Sarahandrano = ergm_full_sar),
  full_term_labels
)

summary(ergm_full_amp)
summary(ergm_full_man)
summary(ergm_full_sar)


meta_df_full <- pool_all(summary_df_full)
print(meta_df_full)

animal_levels_full <- c("Rodent", "Wild",
                        "Vanilla:Rodent",    "Vanilla:Wild",
                        "House Mat.:Rodent", "House Mat.:Wild",
                        "Goods:Rodent",      "Goods:Wild")

# Pooled
meta_full_r <- meta_df_full |> rename_vars()
ggarrange(
  meta_full_r |> filter(Variable %in% people_levels) |>
    coef_plot(people_levels),
  meta_full_r |> filter(Variable %in% animal_levels_full) |>
    coef_plot(animal_levels_full),
  labels = c("a", "b"), common.legend = TRUE
)

# By village
summary_full_r <- summary_df_full |> rename_vars() |>
  rename_villages() |> add_ci_cols("Std. Error")

ggarrange(
  summary_full_r |> filter(Variable %in% people_levels) |>
    coef_plot(people_levels, color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  summary_full_r |> filter(Variable %in% animal_levels_full) |>
    coef_plot(animal_levels_full, color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  labels = c("a", "b"), common.legend = TRUE
)
```

## 6c. High-Risk Network: Pool and Plot

```{r high-risk-pool-plot}
hr_term_labels <- c(
  "Edges", "Age", "Gender[Male]", "CGSOL", "HSOL", "Vanilla",
  "Land Size", "Household Size", "School Level",
  "Other Dom.", "Rodents", "Wild",
  "CGSOL:Other Dom.", "CGSOL:Rodents", "CGSOL:Wild",
  "HSOL:Other Dom.",  "HSOL:Rodents",  "HSOL:Wild",
  "Vanilla:Other Dom.", "Vanilla:Rodents", "Vanilla:Wild"
)

summary_df_hr <- build_summary_df(
  list(Ampandrana = ergm_hr_amp, Mandena = ergm_hr_man,
       Sarahandrano = ergm_hr_sar),
  hr_term_labels
)

summary(ergm_hr_amp)
summary(ergm_hr_man)
summary(ergm_hr_sar)

meta_df_hr <- pool_all(summary_df_hr)
print(meta_df_hr)

animal_levels_hr <- c(
  "Other Dom.", "Rodent", "Wild",
  "Vanilla:Other Dom.", "Vanilla:Rodent",    "Vanilla:Wild",
  "House Mat.:Other Dom.", "House Mat.:Rodent", "House Mat.:Wild",
  "Goods:Other Dom.",   "Goods:Rodent",      "Goods:Wild"
)

# Pooled
meta_hr_r <- meta_df_hr |> rename_vars()
ggarrange(
  meta_hr_r |> filter(Variable %in% people_levels) |>
    coef_plot(people_levels),
  meta_hr_r |> filter(Variable %in% animal_levels_hr) |>
    coef_plot(animal_levels_hr),
  labels = c("a", "b"), common.legend = TRUE
)

# By village
summary_hr_r <- summary_df_hr |> rename_vars() |>
  rename_villages() |> add_ci_cols("Std. Error")

ggarrange(
  summary_hr_r |> filter(Variable %in% people_levels) |>
    coef_plot(people_levels, color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  summary_hr_r |> filter(Variable %in% animal_levels_hr) |>
    coef_plot(animal_levels_hr, color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  labels = c("a", "b"), common.legend = TRUE
)
```

---

# 7. Descriptive Statistics

## 7a. Network-Level Stats Tables

```{r network-stats}
calculate_network_stats <- function(graph, village_name) {
  hd <- V(graph)$degree[ V(graph)$type]
  ad <- V(graph)$degree[!V(graph)$type]
  data.frame(
    Village            = village_name,
    N_Human_Nodes      = sum( V(graph)$type),
    Mean_Human_Degree  = mean(hd), SD_Human_Degree  = sd(hd),
    Min_Human_Degree   = min(hd),  Max_Human_Degree = max(hd),
    N_Animal_Nodes     = sum(!V(graph)$type),
    Mean_Animal_Degree = mean(ad), SD_Animal_Degree  = sd(ad),
    Min_Animal_Degree  = min(ad),  Max_Animal_Degree = max(ad),
    Num_Edges          = igraph::ecount(graph),
    Density            = igraph::edge_density(graph)
  )
}

col_names <- c("Village", "N Human", "Mean Deg.", "SD",
               "Min", "Max", "N Animal", "Mean Ani. Deg.",
               "SD", "Min", "Max", "Edges", "Density")

bind_rows(
  calculate_network_stats(bg_full_man, "Mandena"),
  calculate_network_stats(bg_full_sar, "Sarahandrano"),
  calculate_network_stats(bg_full_amp, "Ampandrana/Andatsakala")
) |> kable(caption = "Full Network — Descriptive Statistics",
           col.names = col_names, digits = 3, row.names = FALSE)

bind_rows(
  calculate_network_stats(bg_hr_man, "Mandena"),
  calculate_network_stats(bg_hr_sar, "Sarahandrano"),
  calculate_network_stats(bg_hr_amp, "Ampandrana/Andatsakala")
) |> kable(caption = "High-Risk Network — Descriptive Statistics",
           col.names = col_names, digits = 3, row.names = FALSE)
```

## 7b. Flat Edge Data Frames

```{r build-flat-edge-dfs}
make_flat_edge_df <- function(merged_df, animals, col_pattern_fn,
                              bipartite_graph) {
  edge_df <- expand.grid(from = unique(merged_df$social_netid),
                          to   = animals,
                          stringsAsFactors = FALSE) |>
    arrange(from) |> mutate(edge = 0L)

  for (animal in animals) {
    cols <- col_pattern_fn(animal, merged_df)
    if (length(cols) == 0) next
    has_contact <- rowSums(as.data.frame(merged_df[cols]),
                           na.rm = TRUE) > 0
    edge_df$edge[edge_df$to == animal &
                   edge_df$from %in%
                   merged_df$social_netid[has_contact]] <- 1L
  }

  vertex_meta <- igraph::as_data_frame(bipartite_graph,
                                       what = "vertices") |>
    rename(from = name) |> filter(animal_name == "Human")

  left_join(edge_df, vertex_meta, by = "from")
}

flat_full <- make_flat_edge_df(
  merged_df, animals_full, full_col_pattern, bipartite_graph_full
)
flat_hr <- make_flat_edge_df(
  merged_df, animals_high_risk,
  function(animal, df) paste0(animal, "_edge_high_risk"),
  bipartite_graph_high_risk
)
```

## 7c. Interaction Concentration and Averages

```{r concentration}
concentration_50 <- function(edge_df, label = "") {
  pct <- edge_df |>
    group_by(from) |>
    summarise(n = sum(edge == 1, na.rm = TRUE), .groups = "drop") |>
    arrange(desc(n)) |>
    mutate(cp = row_number() / n(), ce = cumsum(n) / sum(n)) |>
    filter(ce >= 0.50) |> slice(1) |> pull(cp) * 100
  cat(label, ": ", round(pct, 1),
      "% of people account for 50% of interactions\n", sep = "")
}

avg_interactions <- function(edge_df, label = "") {
  r <- edge_df |> filter(!is.na(edge)) |>
    group_by(from) |>
    summarise(n = sum(edge == 1), .groups = "drop") |>
    summarise(mean = mean(n), sd = sd(n), n = n())
  cat(label, ": mean =", round(r$mean, 2),
      ", SD =", round(r$sd, 2), ", n =", r$n, "\n")
}
# Extract the human vertex IDs that were retained in each statnet network
# (after dropping vertices with missing covariates in igraph_to_statnet())
get_retained_humans <- function(sn) {
  ids <- sn %v% "vertex.names"
  animal_names <- sn %v% "animal_name"
  ids[animal_names == "Human"]
}

# Union of humans retained across all three village ERGMs
full_ergm_humans <- Reduce(union, list(
  get_retained_humans(sn_full_amp),
  get_retained_humans(sn_full_man),
  get_retained_humans(sn_full_sar)
))

hr_ergm_humans <- Reduce(union, list(
  get_retained_humans(sn_hr_amp),
  get_retained_humans(sn_hr_man),
  get_retained_humans(sn_hr_sar)
))

flat_full_ergm <- flat_full |> filter(from %in% full_ergm_humans)
flat_hr_ergm   <- flat_hr   |> filter(from %in% hr_ergm_humans)

concentration_50(flat_full_ergm, "Full network")
concentration_50(flat_hr_ergm,   "High-risk network")
avg_interactions(flat_full_ergm, "Full network")
avg_interactions(flat_hr_ergm,   "High-risk network")
```

## 7d. SES Predictor Comparisons (t-tests)

```{r ses-ttests}
ses_ttest <- function(edge_df, ses_var, binary = FALSE) {
  df <- edge_df |>
    filter(!is.na(edge), !is.na(.data[[ses_var]])) |>
    group_by(from) |>
    summarise(total_edges = sum(edge == 1),
              ses = first(.data[[ses_var]]), .groups = "drop")

  if (binary) {
    t.test(total_edges ~ ses, data = df)
  } else {
    df <- df |>
      mutate(rank = percent_rank(ses),
             quintile = case_when(rank <= 0.20 ~ "bottom_20",
                                  rank >= 0.80 ~ "top_20",
                                  TRUE         ~ NA_character_)) |>
      filter(!is.na(quintile))
    t.test(total_edges ~ quintile, data = df)
  }
}

for (nw in c("full", "hr")) {
  ed <- if (nw == "full") flat_full_ergm else flat_hr_ergm
  cat("\n---", if (nw == "full") "Full" else "High-Risk", "Network ---\n")
  cat("Land size:\n");        print(ses_ttest(ed, "land_size"))
  cat("Commercial goods:\n"); print(ses_ttest(ed, "commercial_goods"))
  cat("House materials:\n");  print(ses_ttest(ed, "house_sol"))
  cat("Vanilla:\n");          print(ses_ttest(ed, "grew_vanilla",
                                              binary = TRUE))
}

```
