---
title: "Self-Reported Network Descriptive Stats"
author: "Lev Kolinski"
date: "`r Sys.Date()`"
output: html_document
---

## Set Up

```{r setup}
library(here)
library(ineq)
library(kableExtra)
library(tidyverse)

library(here)
library(igraph)
source(here("./self_reported_networks/Loading Interaction Data.R"))

# BUILDING FULL NETWORK

# setting animals of interest ----
animals <- c("rodents", "poultry", "dogs", "cats", "domestic_pigs", "goats_sheep",
                  "cows", "wild_birds", "bush_pigs", "lemurs", "tenrecs", "carnivores")

# creating wide df ----
for (animal in animals) {
  # Find columns containing the current animal type
  animal_cols <- grep(paste0("_", animal), names(merged_df), value = TRUE)
  
  # Create edges for animals reported at least once
  merged_df[[paste0(animal, "_edge")]] <- as.integer(rowSums(merged_df[animal_cols]) > 0)
}

animal_bipartite_df_wide <- merged_df |> 
  select(matches("social_netid|_edge", 
                 ignore.case = FALSE)) 

# make wide df into 2-col df----
# Initialize an empty list to store data frames for each animal type
animal_edges_dfs <- list()

for (animal in animals) {
  # Get social_netid values for the current animal type
  animal_edges <- animal_bipartite_df_wide %>%
    filter(!!sym(paste0(animal, "_edge")) == 1) %>%
    select(social_netid) %>%
    as.vector()
  
  # Create a dataframe for the current animal type
  animal_edges_df <- data.frame(social_netid = animal_edges, animal = animal)
  
  # Store the dataframe in the list
  animal_edges_dfs[[animal]] <- animal_edges_df
}

all_animal_edges_df <- bind_rows(animal_edges_dfs)
all_animal_edges_df <- arrange(all_animal_edges_df, social_netid)

# making bipartite graph----
bipartite_graph_full_network <- graph_from_data_frame(all_animal_edges_df, directed = FALSE)


# adding back in social netid isolates
isolates <- setdiff(merged_df$social_netid, all_animal_edges_df$social_netid)
isolate_animals <- setdiff(animals, all_animal_edges_df$animal)

bipartite_graph_full_network <- add_vertices(bipartite_graph_full_network, 
                                          nv = length(isolates), name = isolates)

# Setting vertex types----
animal_vertices <- which(V(bipartite_graph_full_network)$name %in% animals)

V(bipartite_graph_full_network)$type <- TRUE
V(bipartite_graph_full_network)$type[animal_vertices] <- FALSE # animals = FALSE

social_netid_vertices <- which(V(bipartite_graph_full_network)$type) # setting the human mode

## adding in human vector attributes for MI and demographics ----

village_values <- merged_df$village[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
age_values <- merged_df$age[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
gender_values <- merged_df$gender[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
commercial_goods_values <- merged_df$commercial_goods[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
house_sol_values <- merged_df$house_sol[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
grew_vanilla_values <- merged_df$grew_vanilla[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
land_size_values <- merged_df$landsize_in_daba[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
household_size_values <- merged_df$household_size[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]
school_level_values <- merged_df$school_level_numbered[match(V(bipartite_graph_full_network)$name[social_netid_vertices], merged_df$social_netid)]

# Assign the attributes to the social_netid vertices
V(bipartite_graph_full_network)$village <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$village[social_netid_vertices] <- village_values

V(bipartite_graph_full_network)$age <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$age[social_netid_vertices] <- age_values

V(bipartite_graph_full_network)$gender <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$gender[social_netid_vertices] <- gender_values

V(bipartite_graph_full_network)$commercial_goods <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$commercial_goods[social_netid_vertices] <- commercial_goods_values

V(bipartite_graph_full_network)$house_sol <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$house_sol[social_netid_vertices] <- house_sol_values

V(bipartite_graph_full_network)$grew_vanilla <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$grew_vanilla[social_netid_vertices] <- grew_vanilla_values

V(bipartite_graph_full_network)$land_size <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$land_size[social_netid_vertices] <- land_size_values

V(bipartite_graph_full_network)$household_size <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$household_size[social_netid_vertices] <- household_size_values

V(bipartite_graph_full_network)$school_level <- NA  # Initialize the attribute
V(bipartite_graph_full_network)$school_level[social_netid_vertices] <- school_level_values



# adding in animal vector attributes for animal type
V(bipartite_graph_full_network)$animal_name <- "Human"  # Initialize the attribute
V(bipartite_graph_full_network)$animal_name[animal_vertices] <- V(bipartite_graph_full_network)$name[animal_vertices]



# doing different bipartite networks for each village ----
bipartite_graph_full_network_ampandrana_andatsakala <- subgraph(graph = bipartite_graph_full_network, 
                                                                vids = which(V(bipartite_graph_full_network)$village == "Ampandrana" | 
                                                                               V(bipartite_graph_full_network)$village == "Andatsakala" | 
                                                                               is.na(V(bipartite_graph_full_network)$village))) # the na vertices are animals!

V(bipartite_graph_full_network_ampandrana_andatsakala)$degree<-igraph::degree(bipartite_graph_full_network_ampandrana_andatsakala)
mean(V(bipartite_graph_full_network_ampandrana_andatsakala)$degree[V(bipartite_graph_full_network_ampandrana_andatsakala)$type]) ##degrees for humans
range(V(bipartite_graph_full_network_ampandrana_andatsakala)$degree[V(bipartite_graph_full_network_ampandrana_andatsakala)$type]) ##range for humans
mean(V(bipartite_graph_full_network_ampandrana_andatsakala)$degree[!V(bipartite_graph_full_network_ampandrana_andatsakala)$type]) ##degrees for animals
range(V(bipartite_graph_full_network_ampandrana_andatsakala)$degree[!V(bipartite_graph_full_network_ampandrana_andatsakala)$type]) ##range for animals




bipartite_graph_full_network_mandena <- subgraph(graph=bipartite_graph_full_network,
                                                 vids=which(V(bipartite_graph_full_network)$village=="Mandena" |
                                                                                                  is.na(V(bipartite_graph_full_network)$village)))

V(bipartite_graph_full_network_mandena)$degree<-igraph::degree(bipartite_graph_full_network_mandena)
mean(V(bipartite_graph_full_network_mandena)$degree[V(bipartite_graph_full_network_mandena)$type]) ##degrees for humans
range(V(bipartite_graph_full_network_mandena)$degree[V(bipartite_graph_full_network_mandena)$type]) ##range for humans
mean(V(bipartite_graph_full_network_mandena)$degree[!V(bipartite_graph_full_network_mandena)$type]) ##degrees for animals
range(V(bipartite_graph_full_network_mandena)$degree[!V(bipartite_graph_full_network_mandena)$type]) ##range for animals



bipartite_graph_full_network_sarahandrano <- subgraph(graph=bipartite_graph_full_network, 
                                                      vids=which(V(bipartite_graph_full_network)$village=="Sarahandrano" | 
                                                                                                       is.na(V(bipartite_graph_full_network)$village)))
V(bipartite_graph_full_network_sarahandrano)$degree<-igraph::degree(bipartite_graph_full_network_sarahandrano)
mean(V(bipartite_graph_full_network_sarahandrano)$degree[V(bipartite_graph_full_network_sarahandrano)$type]) ##degrees for humans
range(V(bipartite_graph_full_network_sarahandrano)$degree[V(bipartite_graph_full_network_sarahandrano)$type]) ##range for humans
mean(V(bipartite_graph_full_network_sarahandrano)$degree[!V(bipartite_graph_full_network_sarahandrano)$type]) ##degrees for animals
range(V(bipartite_graph_full_network_sarahandrano)$degree[!V(bipartite_graph_full_network_sarahandrano)$type]) ##range for animals


## going back to assigning animals based on domesticated, rodent, or wild

# animal categories
domesticated_animals <- c("poultry", "dogs", "domestic_pigs", "cows", "goats_sheep", "cats")
rodent_animals <- "rodents"
wild_animals <- c("bush_pigs", "carnivores", "tenrecs", "wild_birds", "lemurs")


# Create the new attribute animal_category -----
#ampandrana

V(bipartite_graph_full_network_ampandrana_andatsakala)$animal_category <- ifelse(
  V(bipartite_graph_full_network_ampandrana_andatsakala)$animal_name %in% domesticated_animals, "domesticated",
  ifelse(V(bipartite_graph_full_network_ampandrana_andatsakala)$animal_name %in% wild_animals, "wild", 
         ifelse(V(bipartite_graph_full_network_ampandrana_andatsakala)$animal_name %in% rodent_animals, "rodent", 
        NA )))

#mandena
V(bipartite_graph_full_network_mandena)$animal_category <- ifelse(
  V(bipartite_graph_full_network_mandena)$animal_name %in% domesticated_animals, "domesticated",
  ifelse(V(bipartite_graph_full_network_mandena)$animal_name %in% wild_animals, "wild", 
         ifelse(V(bipartite_graph_full_network_mandena)$animal_name %in% rodent_animals, "rodent", 
                NA )))

#sarahandrano
V(bipartite_graph_full_network_sarahandrano)$animal_category <- ifelse(
  V(bipartite_graph_full_network_sarahandrano)$animal_name %in% domesticated_animals, "domesticated",
  ifelse(V(bipartite_graph_full_network_sarahandrano)$animal_name %in% wild_animals, "wild", 
         ifelse(V(bipartite_graph_full_network_sarahandrano)$animal_name %in% rodent_animals, "rodent", 
                NA )))


# LOADING DFs----
demographic_health_data <- read.csv("/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Demographic_Health.csv")
animal_interaction_data <- read.csv("/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Animal_Interaction.csv")


# cleaning data, selecting only columns that I want, converting binaries to 0 and 1
demographic_health_data_cleaned <- demographic_health_data |> 
  mutate(across(starts_with(c("own_", "banking", "grew_crops",
                              "crops")),
                ~ifelse(. == "Yes", 1L, 0L))) |> 
  select(survey_date, social_netid, village,age,gender,school_level, household_size, main_activity, 
         starts_with("material_"), starts_with("own_"),crops_vanilla, landsize_in_daba)


animal_interaction_data_cleaned <- animal_interaction_data |>  
  mutate(across(-starts_with("social_netid"), ~ifelse(. == "Yes", 1L, 0L))) |> 
  select(-starts_with("symptoms"),-starts_with("action_"), -contains("frequency"))


merged_df <- inner_join(demographic_health_data_cleaned, animal_interaction_data_cleaned, by = "social_netid")


cat("N after inner_join:", nrow(merged_df), "\n")
cat("Unique IDs:", length(unique(merged_df$social_netid)), "\n")

merged_df <- suppressWarnings(merged_df |> 
                                        mutate(material_wall_index = 0) %>%
                                        mutate(material_wall_index = if_else(material_wall == "bamboo", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "rafia", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "ravenala", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "mud", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "compacted_earth", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "wood_planks", 1, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "brick_unfired", 2, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "brick_fired", 2, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "metal_sheets", 3, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "cement", 4, material_wall_index)) %>%
                                        mutate(material_wall_index = scale(material_wall_index)) |> 
                                        
                                        # roof construction
                                        mutate(material_roof_index = 0) %>%
                                        mutate(material_roof_index = if_else(material_roof == "bamboo", 0, material_roof_index)) %>%
                                        mutate(material_roof_index = if_else(material_roof == "thatch", 0, material_roof_index)) %>%
                                        mutate(material_roof_index = if_else(material_roof == "metal_sheets", 1, material_roof_index)) %>%
                                        mutate(material_roof_index = if_else(material_roof == "cement", 2, material_roof_index)) %>%
                                        mutate(material_roof_index = scale(material_roof_index)) |> 
                                        
                                        # floor construction
                                        mutate(material_floor_index = 0) %>%
                                        mutate(material_floor_index = if_else(material_floor == "dirt", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "bamboo", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "rafia", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "ravinala", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "wood_planks", 1, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "cement", 2, material_floor_index)) %>%
                                        mutate(material_floor_index = scale(material_floor_index)) |> 
                                        
                                        # house index
                                        mutate(house_sol = material_wall_index + material_roof_index + material_floor_index) |> 
                                        mutate(house_sol = scale(house_sol)) |> 
                                        
                                        # commercial goods
                                        mutate(commercial_goods = own_cellphone + own_tv + own_bicycle + own_refrigerator +
                                                 own_motorcycle + own_computer + own_generator) |> 
                                        mutate(commercial_goods = scale(commercial_goods)) |> 
                                        
                                        # other things
                                        mutate(employment_category = if_else(grepl("farm_crops", main_activity) | grepl("farm_mixed", main_activity), "Farmer", 
                                                                             ifelse(grepl("student", main_activity) | grepl("teacher", main_activity), "Education",
                                                                                    ifelse(grepl("unemployed", main_activity) | grepl("retired_teacher", main_activity), "Unemployed",
                                                                                           "Other")))) |> 
                                        mutate(school_level_numbered = ifelse(school_level == "None", 0L, 
                                                                              ifelse(school_level == "Primary", 1L,
                                                                                     ifelse(school_level == "Secondary", 2L,
                                                                                            ifelse(school_level == "Higher", 3L,
                                                                                                   NA))))) |> 
                                        rename(grew_vanilla = crops_vanilla) |> 
                                        mutate(age = scale(age)) |> 
                                        mutate(landsize_in_daba = scale(landsize_in_daba)) |> 
                                        mutate(household_size = scale(household_size)) |> 
                                        mutate(school_level_numbered = scale(school_level_numbered))) |> 
  drop_na(house_sol, commercial_goods, grew_vanilla, age, landsize_in_daba, household_size, school_level_numbered, gender)
                                
merged_df$survey_date <- as.Date(merged_df$survey_date)

before_drop <- nrow(merged_df)
dropped_ids <- merged_df |> 
  filter(is.na(house_sol) | is.na(commercial_goods) | is.na(grew_vanilla) | 
         is.na(age) | is.na(landsize_in_daba) | is.na(household_size) | 
         is.na(school_level_numbered) | is.na(gender)) |> 
  pull(social_netid)
cat("N dropped:", length(dropped_ids), "\n")
cat("Dropped IDs:", paste(dropped_ids, collapse = ", "), "\n")

# BUILDING HIGH RISK NETWORK

merged_df<- merged_df |> 
  select(-grep("frequency", names(merged_df)))

# setting animals of interest ----
animals <- c("rodents", "poultry", "dogs", "cats", "domestic_pigs", "goats_sheep",
                  "carnivores", "cows", "wild_birds", "bush_pigs", "lemurs", "tenrecs")
# creating wide df ----
for (animal in animals) {
  # Find columns containing the current animal type
  feces_cols <- grep(paste0("feces_", animal), names(merged_df), value = TRUE)
  shared_water_cols <- grep(paste0("shared_water_", animal), names(merged_df), value = TRUE)
  scratched_bitten_cols <- grep(paste0("scratched_bitten_", animal), names(merged_df), value = TRUE)
  cooked_handled_cols <- grep(paste0("cooked_handled_", animal), names(merged_df), value = TRUE)
  raw_undercooked_cols <- grep(paste0("raw_undercooked_", animal), names(merged_df), value = TRUE)
  eaten_sick_cols <- grep(paste0("eaten_sick_", animal), names(merged_df), value = TRUE)
  slaughtered_cols <- grep(paste0("slaughtered_", animal), names(merged_df), value = TRUE)

  # Create edges for animals reported at least once
  merged_df[[paste0(animal, "_edge_feces")]] <- as.integer(rowSums(merged_df[feces_cols]) > 0)
  merged_df[[paste0(animal, "_edge_shared_water")]] <- as.integer(rowSums(merged_df[shared_water_cols]) > 0)
  merged_df[[paste0(animal, "_edge_scratched_bitten")]] <- as.integer(rowSums(merged_df[scratched_bitten_cols]) > 0)
  merged_df[[paste0(animal, "_edge_cooked_handled")]] <- as.integer(rowSums(merged_df[cooked_handled_cols]) > 0)
  merged_df[[paste0(animal, "_edge_raw_undercooked")]] <- as.integer(rowSums(merged_df[raw_undercooked_cols]) > 0)
  merged_df[[paste0(animal, "_edge_eaten_sick")]] <- as.integer(rowSums(merged_df[eaten_sick_cols]) > 0)
  merged_df[[paste0(animal, "_edge_slaughered")]] <- as.integer(rowSums(merged_df[slaughtered_cols]) > 0)
  
  
  # create high risk edge
  merged_df[[paste0(animal, "_edge_high_risk")]] <- as.integer(rowSums(merged_df[, c(paste0(animal, "_edge_feces"),
                                                                                     paste0(animal, "_edge_shared_water"),
                                                                                     paste0(animal, "_edge_scratched_bitten"),
                                                                                     paste0(animal, "_edge_cooked_handled"),
                                                                                     paste0(animal, "_edge_raw_undercooked"),
                                                                                     paste0(animal, "_edge_eaten_sick"),
                                                                                     paste0(animal, "_edge_slaughered"))]) > 0)
}


animal_bipartite_high_risk_df_wide <- merged_df |> 
  select(matches("social_netid|_edge", 
                 ignore.case = FALSE))

animal_edges_dfs_high_risk <- list()

for (animal in animals) {
  # Get social_netid values for the current animal type
  animal_edges_high_risk <- animal_bipartite_high_risk_df_wide |> 
    filter(!!sym(paste0(animal, "_edge_high_risk")) == 1) |> 
    select(social_netid) |> 
    pull()
  
  # Check if animal_edges_inside is not empty
  if (length(animal_edges_high_risk) > 0) {
    # Create a dataframe for the current animal type
    animal_edges_df_high_risk <- data.frame(social_netid = animal_edges_high_risk, animal = animal)
    
    # Store the dataframe in the list
    animal_edges_dfs_high_risk[[animal]] <- animal_edges_df_high_risk
  }
}
all_animal_edges_df_high_risk <- bind_rows(animal_edges_dfs_high_risk)
all_animal_edges_df_high_risk <- arrange(all_animal_edges_df_high_risk, social_netid)

# making bipartite graph----
bipartite_graph_high_risk <- graph_from_data_frame(all_animal_edges_df_high_risk, directed = FALSE)

# adding back in social netid isolates
isolates <- setdiff(merged_df$social_netid, all_animal_edges_df_high_risk$social_netid)

# adding back in animal isolates
isolate_animals <- setdiff(animals, all_animal_edges_df_high_risk$animal)

bipartite_graph_high_risk <- add_vertices(bipartite_graph_high_risk, 
                                      nv = length(isolates), name = isolates)

bipartite_graph_high_risk <- add_vertices(bipartite_graph_high_risk, 
                                      nv = length(isolate_animals), name = isolate_animals)

# Setting vertex types----
animal_vertices <- which(V(bipartite_graph_high_risk)$name %in% animals)
V(bipartite_graph_high_risk)$type <- TRUE
V(bipartite_graph_high_risk)$type[animal_vertices] <- FALSE # animals = FALSE

social_netid_vertices <- which(V(bipartite_graph_high_risk)$type)

## adding in human vector attributes for MI and demographics ----

village_values <- merged_df$village[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]

age_values <- merged_df$age[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]

gender_values <- merged_df$gender[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]
commercial_goods_values <- merged_df$commercial_goods[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]
house_sol_values <- merged_df$house_sol[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]
grew_vanilla_values <- merged_df$grew_vanilla[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]
land_size_values <- merged_df$landsize_in_daba[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]
household_size_values <- merged_df$household_size[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]
school_level_values <-merged_df$school_level_numbered[match(V(bipartite_graph_high_risk)$name[social_netid_vertices], merged_df$social_netid)]

# Assign the attributes to the social_netid vertices

V(bipartite_graph_high_risk)$village <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$village[social_netid_vertices] <- village_values

V(bipartite_graph_high_risk)$age <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$age[social_netid_vertices] <- age_values

V(bipartite_graph_high_risk)$gender <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$gender[social_netid_vertices] <- gender_values

V(bipartite_graph_high_risk)$commercial_goods <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$commercial_goods[social_netid_vertices] <- commercial_goods_values

V(bipartite_graph_high_risk)$house_sol <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$house_sol[social_netid_vertices] <- house_sol_values

V(bipartite_graph_high_risk)$grew_vanilla <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$grew_vanilla[social_netid_vertices] <- grew_vanilla_values

V(bipartite_graph_high_risk)$land_size <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$land_size[social_netid_vertices] <- land_size_values

V(bipartite_graph_high_risk)$household_size <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$household_size[social_netid_vertices] <- household_size_values

V(bipartite_graph_high_risk)$school_level <- NA  # Initialize the attribute
V(bipartite_graph_high_risk)$school_level[social_netid_vertices] <- school_level_values



# adding in animal vector attributes for animal type
V(bipartite_graph_high_risk)$animal_name <- "Human"  # Initialize the attribute
V(bipartite_graph_high_risk)$animal_name[animal_vertices] <- V(bipartite_graph_high_risk)$name[animal_vertices]


# doing different bipartite networks for each village ----
# ampandrana adatsakala
bipartite_graph_high_risk_ampandrana_andatsakala <- subgraph(graph = bipartite_graph_high_risk, 
                                                                vids = which(V(bipartite_graph_high_risk)$village == "Ampandrana" | 
                                                                               V(bipartite_graph_high_risk)$village == "Andatsakala" | 
                                                                               is.na(V(bipartite_graph_high_risk)$village))) # the na vertices are animals!

V(bipartite_graph_high_risk_ampandrana_andatsakala)$degree<-igraph::degree(bipartite_graph_high_risk_ampandrana_andatsakala)
mean(V(bipartite_graph_high_risk_ampandrana_andatsakala)$degree[V(bipartite_graph_high_risk_ampandrana_andatsakala)$type]) ##degrees for humans
range(V(bipartite_graph_high_risk_ampandrana_andatsakala)$degree[V(bipartite_graph_high_risk_ampandrana_andatsakala)$type]) ##range for humans
mean(V(bipartite_graph_high_risk_ampandrana_andatsakala)$degree[!V(bipartite_graph_high_risk_ampandrana_andatsakala)$type]) ##degrees for animals
range(V(bipartite_graph_high_risk_ampandrana_andatsakala)$degree[!V(bipartite_graph_high_risk_ampandrana_andatsakala)$type]) ##range for animals


# mandena
bipartite_graph_high_risk_mandena <- subgraph(graph=bipartite_graph_high_risk,
                                                 vids=which(V(bipartite_graph_high_risk)$village=="Mandena" |
                                                              is.na(V(bipartite_graph_high_risk)$village)))


V(bipartite_graph_high_risk_mandena)$degree<-igraph::degree(bipartite_graph_high_risk_mandena)
mean(V(bipartite_graph_high_risk_mandena)$degree[V(bipartite_graph_high_risk_mandena)$type]) ##degrees for humans
range(V(bipartite_graph_high_risk_mandena)$degree[V(bipartite_graph_high_risk_mandena)$type]) ##range for humans
mean(V(bipartite_graph_high_risk_mandena)$degree[!V(bipartite_graph_high_risk_mandena)$type]) ##degrees for animals
range(V(bipartite_graph_high_risk_mandena)$degree[!V(bipartite_graph_high_risk_mandena)$type]) ##range for animals

# sarahandrano
bipartite_graph_high_risk_sarahandrano <- subgraph(graph=bipartite_graph_high_risk, 
                                                      vids=which(V(bipartite_graph_high_risk)$village=="Sarahandrano" | 
                                                                   is.na(V(bipartite_graph_high_risk)$village)))

V(bipartite_graph_high_risk_sarahandrano)$degree<-igraph::degree(bipartite_graph_high_risk_sarahandrano)
mean(V(bipartite_graph_high_risk_sarahandrano)$degree[V(bipartite_graph_high_risk_sarahandrano)$type]) ##degrees for humans
range(V(bipartite_graph_high_risk_sarahandrano)$degree[V(bipartite_graph_high_risk_sarahandrano)$type]) ##range for humans
mean(V(bipartite_graph_high_risk_sarahandrano)$degree[!V(bipartite_graph_high_risk_sarahandrano)$type]) ##degrees for animals
range(V(bipartite_graph_high_risk_sarahandrano)$degree[!V(bipartite_graph_high_risk_sarahandrano)$type]) ##range for animals


# animal categories
domesticated_animals_1 <- c("poultry", "domestic_pigs", "cows")
domesticated_animals_2 <- c( "dogs","goats_sheep", "cats")
rodent_animals <- "rodents"
wild_animals <- c("bush_pigs", "carnivores", "tenrecs", "wild_birds", "lemurs")

# domesticated_animals <- c("poultry", "domestic_pigs", "cows", "dogs","goats_sheep", "cats")

# Create the new attribute animal_category -----
#ampandrana
V(bipartite_graph_high_risk_ampandrana_andatsakala)$animal_category <- ifelse(
  V(bipartite_graph_high_risk_ampandrana_andatsakala)$animal_name %in% domesticated_animals_1, "domesticated_1",
  ifelse( V(bipartite_graph_high_risk_ampandrana_andatsakala)$animal_name %in% domesticated_animals_2, "domesticated_2",
  ifelse(V(bipartite_graph_high_risk_ampandrana_andatsakala)$animal_name %in% wild_animals, "wild", 
         ifelse(V(bipartite_graph_high_risk_ampandrana_andatsakala)$animal_name %in% rodent_animals, "rodent", 
                NA ))))

#mandena
V(bipartite_graph_high_risk_mandena)$animal_category <- ifelse(
  V(bipartite_graph_high_risk_mandena)$animal_name %in% domesticated_animals_1, "domesticated_1",
  ifelse( V(bipartite_graph_high_risk_mandena)$animal_name %in% domesticated_animals_2, "domesticated_2",
          ifelse(V(bipartite_graph_high_risk_mandena)$animal_name %in% wild_animals, "wild", 
                 ifelse(V(bipartite_graph_high_risk_mandena)$animal_name %in% rodent_animals, "rodent", 
                        NA ))))

#sarahandrano
V(bipartite_graph_high_risk_sarahandrano)$animal_category <- ifelse(
  V(bipartite_graph_high_risk_sarahandrano)$animal_name %in% domesticated_animals_1, "domesticated_1",
  ifelse( V(bipartite_graph_high_risk_sarahandrano)$animal_name %in% domesticated_animals_2, "domesticated_2",
          ifelse(V(bipartite_graph_high_risk_sarahandrano)$animal_name %in% wild_animals, "wild", 
                 ifelse(V(bipartite_graph_high_risk_sarahandrano)$animal_name %in% rodent_animals, "rodent", 
                        NA ))))



```

## FULL NETWORK

### Network Descriptive Stats Table 
```{r}

# Function to calculate degree statistics and network properties for each village
calculate_network_stats <- function(graph, village_name) {
  human_degrees <- V(graph)$degree[V(graph)$type]  # Degrees for humans
  animal_degrees <- V(graph)$degree[!V(graph)$type]  # Degrees for animals
  
  
  # Collect statistics
  stats <- data.frame(
    Village = village_name,
    Number_Human_Nodes = sum(V(graph)$type == TRUE),
    Mean_Human_Degree = mean(human_degrees),
    SD_Human_Degree = sd(human_degrees),
    Min_Human_Degree = min(human_degrees),
    Max_Human_Degree = max(human_degrees),
    Number_Animal_Nodes = sum(V(graph)$type == FALSE),
    Mean_Animal_Degree = mean(animal_degrees),
    SD_Animal_Degree = sd(animal_degrees),
    Min_Animal_Degree = min(animal_degrees),
    Max_Animal_Degree = max(animal_degrees),
    Num_Edges = ecount(graph),
    Density = igraph::edge_density(graph)
  )
  
  return(stats)
}

# List of graph objects and corresponding village names
graphs <- list(
  "Ampandrana Andatsakala" = bipartite_graph_full_network_ampandrana_andatsakala,
  "Mandena" = bipartite_graph_full_network_mandena,
  "Sarahandrano" = bipartite_graph_full_network_sarahandrano
)

# Iterate over graphs and village names to calculate stats
village_stats_list <- lapply(names(graphs), function(village_name) {
  calculate_network_stats(graphs[[village_name]], village_name)
})

# Combine the results into a single data frame
village_stats <- bind_rows(village_stats_list)
village_stats$Village <- factor(village_stats$Village, levels = c("Mandena","Sarahandrano", "Ampandrana Andatsakala"))
village_stats <- village_stats[order(village_stats$Village), ]


# Print the results as a table using knitr::kable for clean display in R Markdown
kable(village_stats, 
      caption = "Full Network Statistics Comparison Across Villages",
      col.names = c("Village","Number Human Nodes", "Mean Human Degree", "SD Human Degree", "Min Human Degree", 
                    "Max Human Degree", "Number Animal Nodes", "Mean Animal Degree", "SD Animal Degree", "Min Animal Degree",
                    "Max Animal Degree",
                    "Number of Edges", "Density"),
      digits = 4,
      row.names = FALSE) 
```

```{r}

# FULL NETWORKS ----

### get all unique animals
animals <- c("rodents", "poultry", "dogs", "cats", "domestic_pigs", "goats_sheep",
             "cows", "wild_birds", "bush_pigs", "lemurs", "tenrecs", "carnivores")

# Get all unique humans from merged survey df
humans <- unique(merged_df$social_netid)

# Create all possible combinations of humans and animals
edge_df <- expand.grid(from = humans, to = animals, stringsAsFactors = FALSE)
edge_df<- edge_df |> 
  arrange(from)

# Initialize edge column to 0
edge_df$edge <- 0

# Loop over each animal type
for (animal in animals) {
  
  animal_cols <- colnames(merged_df)[grepl(animal, colnames(merged_df))]
  
  if (length(animal_cols) > 0) {
    # Check if any contact occurred for each person
    merged_df[[paste0("has_", animal, "_contact")]] <- rowSums(merged_df[animal_cols], na.rm = TRUE) > 0
    
    # Update the edge_df accordingly
    edge_df$edge[edge_df$to == animal & 
                   edge_df$from %in% merged_df$social_netid[merged_df[[paste0("has_", animal, "_contact")]]]] <- 1
  }
}

# adding in animal_category
domesticated_animals <- c("poultry", "dogs", "domestic_pigs", "cows", "goats_sheep", "cats")
rodent_animals <- "rodents"
wild_animals <- c("bush_pigs", "carnivores", "tenrecs", "wild_birds", "lemurs")

edge_df<-edge_df |> 
  mutate(animal_category = ifelse(to %in% domesticated_animals, "domesticated",
                                  ifelse(to %in% rodent_animals, "rodent",
                                         ifelse(to%in%wild_animals, "wild", NA))))

# bring in human survey data
bipartite_graph_full_network_meta_data<-igraph::as_data_frame(bipartite_graph_full_network ,
                                                              what = "vertices") |> 
  rename(from = name)

# merge with edge list
full_network_df<-merge(edge_df,bipartite_graph_full_network_meta_data,
                       by = "from")
```

### Only X% of Individuals Contribute 50% of All INTERACTIONS
```{r}
full_network_df |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  arrange(desc(total_edges)) |> 
  mutate(
    cum_people = row_number() / n(),
    cum_edges  = cumsum(total_edges) / sum(total_edges)
  ) |> 
  filter(cum_edges >= 0.50) |> 
  slice(1) |> 
  summarise(
    percent_of_people_for_half_interactions = 100 * cum_people
  )
```

### Average number interactions per person
```{r}
full_network_df |> 
  filter(!is.na(edge)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    .groups = "drop"
  ) |> 
  summarise(
    average_interactions_per_person = mean(total_edges),
    sd_interactions = sd(total_edges),
    n_people = n()
  )
```

### Land Size Differences
```{r}
landsize_quintile_edges <- full_network_df |> 
  filter(!is.na(edge), !is.na(land_size)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    land_size = first(land_size),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(land_size),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = landsize_quintile_edges)
```

### COMMERCIAL GOODS DIFFERENCES 
```{r}
commercial_goods_quintile_edges <- full_network_df |> 
  filter(!is.na(edge), !is.na(commercial_goods)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    commercial_goods = first(commercial_goods),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(commercial_goods),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = commercial_goods_quintile_edges)
```


### HOUSE MATERIALS DIFFERENCES 
```{r}
house_mat_quintile_edges <- full_network_df |> 
  filter(!is.na(edge), !is.na(house_sol)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    house_sol = first(house_sol),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(house_sol),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = house_mat_quintile_edges)
```

### VANILLA DIFFERNCES
```{r}
edge_summary_vanilla <- full_network_df |> 
  filter(!is.na(edge), !is.na(grew_vanilla)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    grew_vanilla = first(grew_vanilla),
    .groups = "drop"
  )

t.test(total_edges ~ grew_vanilla, data = edge_summary_vanilla)
```

## High-Risk Network

### Network Descriptive Stats Table 
```{r}
graphs_high_risk <- list(
  "Ampandrana Andatsakala" = bipartite_graph_high_risk_ampandrana_andatsakala,
  "Mandena" = bipartite_graph_high_risk_mandena,
  "Sarahandrano" = bipartite_graph_high_risk_sarahandrano
)

# Iterate over graphs_high_risk and village names to calculate stats
village_stats_list_high_risk <- lapply(names(graphs_high_risk), function(village_name) {
  calculate_network_stats(graphs_high_risk[[village_name]], village_name)
})

# Combine the results into a single data frame
village_stats_high_risk <- bind_rows(village_stats_list_high_risk)
village_stats_high_risk$Village <- factor(village_stats_high_risk$Village, levels = c("Mandena","Sarahandrano", "Ampandrana Andatsakala"))
village_stats_high_risk <- village_stats_high_risk[order(village_stats_high_risk$Village), ]

# Print the results as a table using knitr::kable for clean display in R Markdown
kable(village_stats_high_risk, 
      caption = "Full Network Statistics Comparison Across Villages",
      col.names = c("Village", "Number Human Nodes", "Mean Human Degree", "SD Human Degree", "Min Human Degree", 
                    "Max Human Degree", "Number Animal Nodes", "Mean Animal Degree", "SD Animal Degree", "Min Animal Degree",
                    "Max Animal Degree",
                    "Number of Edges", "Density"),
      digits = 4,
      row.names = FALSE)

```


```{r}

# get all unique animals
animals <- c("rodents", "poultry", "dogs", "cats", "domestic_pigs", "goats_sheep",
             "cows", "wild_birds", "bush_pigs", "lemurs", "tenrecs", "carnivores")

# Get all unique humans from merged survey df
humans <- unique(merged_df$social_netid)

# Create all possible combinations of humans and animals
edge_df_high_risk <- expand.grid(from = humans, to = animals, stringsAsFactors = FALSE)
edge_df_high_risk<- edge_df_high_risk |> 
  arrange(from)

# Initialize edge column to 0
edge_df_high_risk$edge <- 0

# Loop over each animal type
for (animal in animals) {
  
  animal_cols <- colnames(merged_df)[grepl(animal, colnames(merged_df))&grepl("feces|shared_water|scratched|cooked|raw|sick|slaughtered", 
                                                                              colnames(merged_df))]
  
  if (length(animal_cols) > 0) {
    # Check if any contact occurred for each person
    merged_df[[paste0("has_", animal, "_contact")]] <- rowSums(merged_df[animal_cols], na.rm = TRUE) > 0
    
    # Update the edge_df accordingly
    edge_df_high_risk$edge[edge_df_high_risk$to == animal & 
                             edge_df_high_risk$from %in% merged_df$social_netid[merged_df[[paste0("has_", animal, "_contact")]]]] <- 1
  }
}



# adding in animal_category
domesticated_animals_1 <- c("poultry", "domestic_pigs", "cows")
domesticated_animals_2 <- c( "dogs","goats_sheep", "cats")
rodent_animals <- "rodents"
wild_animals <- c("bush_pigs", "carnivores", "tenrecs", "wild_birds", "lemurs")

edge_df_high_risk<-edge_df_high_risk |> 
  mutate(animal_category = ifelse(to %in% domesticated_animals_1, "domesticated_1",
                                  ifelse(to %in% domesticated_animals_2, "domesticated_2",
                                         ifelse(to%in%rodent_animals, "rodent", 
                                                ifelse(to%in%wild_animals, "wild", NA)))))

# bring in human survey data
bipartite_graph_high_risk_network_meta_data<-igraph::as_data_frame(bipartite_graph_high_risk ,
                                                                   what = "vertices") |> 
  rename(from = name)

# merge with edge list
high_risk_network_df<-merge(edge_df_high_risk,bipartite_graph_high_risk_network_meta_data,
                            by = "from")

```


### Only X% of Individuals Contribute 50% of All INTERACTIONS
```{r}
high_risk_network_df |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  arrange(desc(total_edges)) |> 
  mutate(
    cum_people = row_number() / n(),
    cum_edges  = cumsum(total_edges) / sum(total_edges)
  ) |> 
  filter(cum_edges >= 0.50) |> 
  slice(1) |> 
  summarise(
    percent_of_people_for_half_interactions = 100 * cum_people
  )

```

### Average number interactions per person
```{r}
high_risk_network_df |> 
  filter(!is.na(edge)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    .groups = "drop"
  ) |> 
  summarise(
    average_interactions_per_person = mean(total_edges),
    sd_interactions = sd(total_edges),
    n_people = n()
  )

```


### LAND SIZE DIFFERENCES
```{r}
landsize_quintile_edges <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(land_size)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    land_size = first(land_size),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(land_size),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = landsize_quintile_edges)
```

### COMMERCIAL GOODS DIFFERENCES
```{r}
commercial_goods_quintile_edges <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(commercial_goods)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    commercial_goods = first(commercial_goods),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(commercial_goods),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = commercial_goods_quintile_edges)
```

### HOUSE MATERIALS DIFFERENCES 
```{r}
house_mat_quintile_edges <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(house_sol)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    house_sol = first(house_sol),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(house_sol),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = house_mat_quintile_edges)

```

### VANILLA DIFFERNCES
```{r}
edge_summary_vanilla <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(grew_vanilla)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    grew_vanilla = first(grew_vanilla),
    .groups = "drop"
  )

t.test(total_edges ~ grew_vanilla, data = edge_summary_vanilla)
```


