---
title: "Supplementary Analysis: Multilevel Logistic Regression"
author: "Lev Kolinski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(igraph)
library(glmmTMB)
library(broom.mixed)
library(metafor)
library(ggpubr)
library(kableExtra)

set.seed(1234)
```

---

# 1. Load Raw Data

```{r load-data}
demographic_health_data <- read.csv(
  "/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Demographic_Health.csv"
)
animal_interaction_data <- read.csv(
  "/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Animal_Interaction.csv"
)
```

---

# 2. Clean and Merge Data

```{r clean-merge}
demographic_health_data_cleaned <- demographic_health_data |>
  mutate(across(
    starts_with(c("own_", "banking", "grew_crops", "crops")),
    ~ ifelse(. == "Yes", 1L, 0L)
  )) |>
  select(
    survey_date, social_netid, village, age, gender, school_level,
    household_size, main_activity,
    starts_with("material_"), starts_with("own_"),
    crops_vanilla, landsize_in_daba
  )

animal_interaction_data_cleaned <- animal_interaction_data |>
  mutate(across(-starts_with("social_netid"), ~ ifelse(. == "Yes", 1L, 0L))) |>
  select(-starts_with("symptoms"), -starts_with("action_"),
         -contains("frequency"))

merged_df_raw <- inner_join(
  demographic_health_data_cleaned,
  animal_interaction_data_cleaned,
  by = "social_netid"
)

cat("N after inner_join:", nrow(merged_df_raw), "\n")
cat("Unique IDs:", length(unique(merged_df_raw$social_netid)), "\n")
```

---

# 3. Feature Engineering

```{r feature-functions}
safe_scale <- function(x, label = "") {
  s <- sd(x, na.rm = TRUE)
  if (is.na(s) || s == 0) {
    if (nchar(label) > 0)
      message("safe_scale: zero/NA variance in '", label,
              "' - returning 0 for non-NA rows")
    return(ifelse(is.na(x), NA_real_, 0))
  }
  as.numeric(scale(x))
}

 
prepare_village_data <- function(df, village_filter = NULL) {

  if (!is.null(village_filter)) {
    df <- df |> filter(village %in% village_filter)
  }

  vname <- paste(village_filter, collapse = "/")

  df <- df |>
    mutate(
      material_wall_index = case_when(
        material_wall %in% c("bamboo", "rafia", "ravenala",
                             "mud", "compacted_earth") ~ 0,
        material_wall == "wood_planks"                 ~ 1,
        material_wall %in% c("brick_unfired",
                             "brick_fired")            ~ 2,
        material_wall == "metal_sheets"                ~ 3,
        material_wall == "cement"                      ~ 4,
        TRUE                                           ~ NA_real_
      ),
      material_roof_index = case_when(
        material_roof %in% c("bamboo", "thatch") ~ 0,
        material_roof == "wood_planks"            ~ 0.5,
        material_roof == "metal_sheets"           ~ 1,
        material_roof == "cement"                 ~ 2,
        TRUE                                      ~ NA_real_
      ),
      material_floor_index = case_when(
        material_floor %in% c("dirt", "bamboo",
                              "rafia", "ravenala") ~ 0,
        material_floor == "wood_planks"            ~ 1,
        material_floor == "cement"                 ~ 2,
        TRUE                                       ~ NA_real_
      )
    )

  df |>
    mutate(
      house_sol = safe_scale(
        material_wall_index + material_roof_index + material_floor_index,
        paste(vname, "house_sol")
      ),
      commercial_goods = safe_scale(
        own_cellphone + own_tv + own_bicycle + own_refrigerator +
          own_motorcycle + own_computer + own_generator,
        paste(vname, "commercial_goods")
      ),
      total_animal_interactions = rowSums(
        across(pet_dogs:dead_goats_sheep), na.rm = TRUE
      ),
      employment_category = case_when(
        grepl("farm_crops|farm_mixed", main_activity)      ~ "Farmer",
        grepl("student|teacher", main_activity)            ~ "Education",
        grepl("unemployed|retired_teacher", main_activity) ~ "Unemployed",
        TRUE                                               ~ "Other"
      ),
      num_high_risk_exposures = rowSums(
        across(contains(c("shared_water", "scratched_bitten", "feces",
                          "dead", "raw_undercooked", "eaten_sick", "slaughtered"))),
        na.rm = TRUE
      ),
      high_risk_exposures_binary = ifelse(
        num_high_risk_exposures == 0L, 0L, 1L
      ),
      school_level_numbered = safe_scale(
        case_when(
          school_level == "None"      ~ 0L,
          school_level == "Primary"   ~ 1L,
          school_level == "Secondary" ~ 2L,
          school_level == "Higher"    ~ 3L,
          TRUE                        ~ NA_integer_
        ),
        paste(vname, "school_level")
      ),
      age              = safe_scale(age,              paste(vname, "age")),
      landsize_in_daba = safe_scale(landsize_in_daba, paste(vname, "land_size")),
      household_size   = safe_scale(household_size,   paste(vname, "household_size"))
    ) |>
    rename(grew_vanilla = crops_vanilla)
}
```

```{r apply-features}
merged_df <- bind_rows(
  prepare_village_data(merged_df_raw, village_filter = c("Ampandrana", "Andatsakala")),
  prepare_village_data(merged_df_raw, village_filter = "Mandena"),
  prepare_village_data(merged_df_raw, village_filter = "Sarahandrano")
) |>
  mutate(survey_date = as.Date(survey_date))

cat("Total respondents:", length(unique(merged_df$social_netid)), "\n")
```

---

# 4. Animal Lists and Categories

```{r animal-lists}
animals_full      <- c("rodents", "poultry", "dogs", "cats",
                       "domestic_pigs", "goats_sheep", "cows",
                       "wild_birds", "bush_pigs", "lemurs",
                       "tenrecs", "carnivores")
animals_high_risk <- animals_full

# Full network: 3 animal categories
domesticated_animals <- c("poultry", "dogs", "domestic_pigs",
                           "cows", "goats_sheep", "cats")
rodent_animals       <- "rodents"
wild_animals         <- c("bush_pigs", "carnivores", "tenrecs",
                           "wild_birds", "lemurs")

# High-risk network: domesticated split into two groups
domesticated_animals_1 <- c("poultry", "domestic_pigs", "cows")
domesticated_animals_2 <- c("dogs", "goats_sheep", "cats")

village_groups <- list(
  ampandrana_andatsakala = c("Ampandrana", "Andatsakala"),
  mandena                = "Mandena",
  sarahandrano           = "Sarahandrano"
)
```

---

# 5. Build Flat Edge Data Frames

```{r col-pattern-functions}
full_col_pattern <- function(animal, df) {
  grep(paste0("_", animal, "$|_", animal, "_"), names(df), value = TRUE)
}

high_risk_col_pattern <- function(animal, df) {
  grep(
    paste0("(feces|shared_water|scratched_bitten|cooked_handled|",
           "raw_undercooked|eaten_sick|slaughtered)_", animal),
    names(df), value = TRUE
  )
}
```

```{r build-high-risk-cols}
# Create binary high-risk edge columns on merged_df
for (animal in animals_high_risk) {
  cols <- high_risk_col_pattern(animal, merged_df)
  if (length(cols) == 0) {
    warning("No high-risk columns found for: ", animal)
    merged_df[[paste0(animal, "_edge_high_risk")]] <- 0L
  } else {
    merged_df[[paste0(animal, "_edge_high_risk")]] <-
      as.integer(rowSums(merged_df[cols], na.rm = TRUE) > 0)
  }
}
```

```{r build-flat-edge-dfs}
# ------------------------------------------------------------------
# make_flat_edge_df(): builds a long-format data frame with one row
# per person-animal dyad, with all human-level covariates attached.
# This is the input format for the MLMs.
# ------------------------------------------------------------------
make_flat_edge_df <- function(df, animals, col_pattern_fn) {
  edge_df <- expand.grid(
    from = unique(df$social_netid),
    to   = animals,
    stringsAsFactors = FALSE
  ) |>
    arrange(from) |>
    mutate(edge = 0L)

  for (animal in animals) {
    cols <- col_pattern_fn(animal, df)
    if (length(cols) == 0) next
    has_contact <- rowSums(as.data.frame(df[cols]), na.rm = TRUE) > 0
    edge_df$edge[edge_df$to == animal &
                   edge_df$from %in% df$social_netid[has_contact]] <- 1L
  }

  # Attach human-level covariates
  human_meta <- df |>
    select(social_netid, village, age, gender, commercial_goods,
           house_sol, grew_vanilla, landsize_in_daba, household_size,
           school_level_numbered) |>
    rename(
      from         = social_netid,
      land_size    = landsize_in_daba,
      school_level = school_level_numbered
    )

  left_join(edge_df, human_meta, by = "from")
}

flat_full <- make_flat_edge_df(merged_df, animals_full, full_col_pattern)
flat_hr   <- make_flat_edge_df(
  merged_df, animals_high_risk,
  function(animal, df) paste0(animal, "_edge_high_risk")
)
```

---

# 6. Identify Participants Retained After NA Removal

This mirrors the vertex-dropping step in `igraph_to_statnet()` from the main
pipeline, ensuring the MLM sample matches the ERGM sample exactly.

```{r retained-humans}
covariate_cols <- c("age", "gender", "commercial_goods", "house_sol",
                    "grew_vanilla", "landsize_in_daba", "household_size",
                    "school_level_numbered")

complete_humans <- merged_df |>
  filter(if_all(all_of(covariate_cols), ~ !is.na(.))) |>
  pull(social_netid) |>
  unique()

cat("Participants with complete covariates:", length(complete_humans), "\n")
cat("Participants dropped due to missing covariates:",
    length(unique(merged_df$social_netid)) - length(complete_humans), "\n")

flat_full_ergm <- flat_full |> filter(from %in% complete_humans)
flat_hr_ergm   <- flat_hr   |> filter(from %in% complete_humans)
```

---

# 7. Prepare Dyad-Level Data for MLMs

```{r prep-dyad-data}
# Full network: 3 animal categories, reference = domesticated
flat_full_ergm <- flat_full_ergm |>
  mutate(
    animal_category = case_when(
      to %in% domesticated_animals ~ "domesticated",
      to %in% rodent_animals       ~ "rodent",
      to %in% wild_animals         ~ "wild",
      TRUE                         ~ NA_character_
    ),
    animal_category = relevel(factor(animal_category), ref = "domesticated"),
    gender          = factor(gender),
    grew_vanilla    = factor(grew_vanilla)
  )

# High-risk network: 4 animal categories, reference = domesticated_1
flat_hr_ergm <- flat_hr_ergm |>
  mutate(
    animal_category = case_when(
      to %in% domesticated_animals_1 ~ "domesticated_1",
      to %in% domesticated_animals_2 ~ "domesticated_2",
      to %in% rodent_animals         ~ "rodent",
      to %in% wild_animals           ~ "wild",
      TRUE                           ~ NA_character_
    ),
    animal_category = relevel(factor(animal_category), ref = "domesticated_1"),
    gender          = factor(gender),
    grew_vanilla    = factor(grew_vanilla)
  )
```

---

# 8. Fit Multilevel Logistic Regression Models

## 8a. Model-Fitting Helper

```{r mlm-helper}
# ------------------------------------------------------------------
# run_mlm(): fits a mixed-effects logistic regression for one village
# subset using glmmTMB. Participant ID (from) is included as a random
# intercept to account for non-independence of dyads from the same
# individual. Animal category interactions with commercial_goods,
# house_sol, and grew_vanilla mirror the b1factor():b2cov() /
# b1factor():b2factor() ERGM terms from the main analysis.
# glmmTMB is preferred over lme4::glmer for better convergence and
# more robust SE estimation with complex interaction structures.
# ------------------------------------------------------------------
run_mlm <- function(df, village_names, label) {
  d <- df |>
    filter(village %in% village_names) |>
    filter(!is.na(animal_category))

  fit <- glmmTMB(
    edge ~
      age + gender + commercial_goods + house_sol +
      grew_vanilla + land_size + household_size + school_level +
      animal_category +
      animal_category:commercial_goods +
      animal_category:house_sol +
      animal_category:grew_vanilla +
      (1 | from),
    data   = d,
    family = binomial
  )
  
  print(AIC(fit))

  tidy(fit, conf.int = TRUE, exponentiate = FALSE) |>
    filter(effect == "fixed") |>
    mutate(Village = label)
}
```

## 8b. Full Network Models

```{r mlm-full, cache=FALSE, message=TRUE, warning=FALSE}
mlm_full_amp <- run_mlm(flat_full_ergm, c("Ampandrana", "Andatsakala"), "Ampandrana")
mlm_full_man <- run_mlm(flat_full_ergm, "Mandena",                       "Mandena")
mlm_full_sar <- run_mlm(flat_full_ergm, "Sarahandrano",                  "Sarahandrano")

mlm_full_summary <- bind_rows(mlm_full_amp, mlm_full_man, mlm_full_sar)
```

## 8c. High-Risk Network Models

```{r mlm-hr, cache=FALSE, message=TRUE, warning=FALSE}
mlm_hr_amp <- run_mlm(flat_hr_ergm, c("Ampandrana", "Andatsakala"), "Ampandrana")
mlm_hr_man <- run_mlm(flat_hr_ergm, "Mandena",                       "Mandena")
mlm_hr_sar <- run_mlm(flat_hr_ergm, "Sarahandrano",                  "Sarahandrano")

mlm_hr_summary <- bind_rows(mlm_hr_amp, mlm_hr_man, mlm_hr_sar)
```

---

# 9. Pool Effect Sizes via Meta-Analysis

```{r pool-helper}
# ------------------------------------------------------------------
# pool_mlm(): fixed-effects meta-analysis across villages.
# Identical FE approach to the ERGM pooling in the main analysis.
# ------------------------------------------------------------------
pool_mlm <- function(summary_df, drop_intercept = TRUE) {
  df <- summary_df
  if (drop_intercept) df <- df |> filter(term != "(Intercept)")

  df |>
    group_by(term) |>
    group_modify(~ {
      fit <- rma(yi = .x$estimate, sei = .x$std.error, method = "FE")
      data.frame(
        Estimate = as.numeric(fit$beta),
        SE       = fit$se,
        Lower    = fit$ci.lb,
        Upper    = fit$ci.ub,
        p        = fit$pval,
        N        = nrow(.x)
      )
    }) |>
    ungroup() |>
    mutate(
      lower_ci    = Estimate - 1.96        * SE,
      upper_ci    = Estimate + 1.96        * SE,
      lower_ci_90 = Estimate - qnorm(0.95) * SE,
      upper_ci_90 = Estimate + qnorm(0.95) * SE
    )
}
```

```{r pool-results}
meta_mlm_full <- pool_mlm(mlm_full_summary)
meta_mlm_hr   <- pool_mlm(mlm_hr_summary)

print(meta_mlm_full)
print(meta_mlm_hr)
```

---

# 10. Plots

## 10a. Plotting Helper

```{r plot-helper}
# ------------------------------------------------------------------
# rename_mlm_terms(): cleans raw glmmTMB term names into the same
# human-readable labels used in the ERGM plots in the main analysis.
# ------------------------------------------------------------------
rename_mlm_terms <- function(df) {
  df |>
    mutate(term = term |>
      str_replace_all("animal_categorydomesticated_2", "Other Dom.") |>
      str_replace_all("animal_categoryrodent",         "Rodent") |>
      str_replace_all("animal_categorywild",           "Wild") |>
      str_replace_all("commercial_goods",              "Goods") |>
      str_replace_all("house_sol",                     "House Mat.") |>
      str_replace_all("grew_vanilla1",                 "Vanilla") |>
      str_replace_all("genderMale",                    "Gender[Man]") |>
      str_replace_all("land_size",                     "Land Size") |>
      str_replace_all("household_size",                "Household Size") |>
      str_replace_all("school_level",                  "School Level") |>
      str_replace_all("^age$",                         "Age") |>
      # Clean up interaction separators
      str_replace_all(":",                             ":")
    )
}

mlm_coef_plot <- function(df, var_levels, title = NULL, color_var = NULL) {
  d <- df |>
    filter(term %in% var_levels) |>
    mutate(term = factor(term, levels = rev(var_levels)))

  aes_base <- if (!is.null(color_var))
    aes(x = Estimate, y = term, color = .data[[color_var]])
  else
    aes(x = Estimate, y = term)

  ggplot(d, aes_base) +
    geom_point(position = position_dodge(0.5), size = 3) +
    geom_linerange(aes(xmin = lower_ci, xmax = upper_ci),
                   position = position_dodge(0.5), linewidth = 0.4) +
    geom_linerange(aes(xmin = lower_ci_90, xmax = upper_ci_90),
                   position = position_dodge(0.5), linewidth = 1.2) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
    labs(x = "Estimated Effect", y = NULL, title = title) +
    theme_classic() +
    theme(legend.position = "top", axis.title.x = element_text(size = 12))
}

# Explicit term orderings matching the ERGM plots exactly
# Full network (3 animal categories)
people_levels_full <- c("House Mat.", "Goods", "Vanilla", "Land Size",
                        "Household Size", "School Level", "Gender[Man]", "Age")
animal_levels_full <- c("Rodent", "Wild",
                        "Vanilla:Rodent",    "Vanilla:Wild",
                        "House Mat.:Rodent", "House Mat.:Wild",
                        "Goods:Rodent",      "Goods:Wild")

# High-risk network (4 animal categories, adds Other Dom.)
people_levels_hr <- c("House Mat.", "Goods", "Vanilla", "Land Size",
                      "Household Size", "School Level", "Gender[Man]", "Age")
animal_levels_hr <- c("Other Dom.", "Rodent", "Wild",
                      "Vanilla:Other Dom.", "Vanilla:Rodent",    "Vanilla:Wild",
                      "House Mat.:Other Dom.", "House Mat.:Rodent", "House Mat.:Wild",
                      "Goods:Other Dom.",      "Goods:Rodent",      "Goods:Wild")
```

## 10b. Full Network - Pooled

```{r plot-full-pooled, fig.width=10, fig.height=6}
meta_mlm_full_r <- meta_mlm_full |> rename_mlm_terms()

ggarrange(
  mlm_coef_plot(meta_mlm_full_r, people_levels_full, title = "Human Attributes"),
  mlm_coef_plot(meta_mlm_full_r, animal_levels_full, title = "Animal Attributes"),
  labels = c("a", "b"), common.legend = TRUE
)
```

## 10c. Full Network - By Village

```{r plot-full-village, fig.width=10, fig.height=6}
mlm_full_village <- mlm_full_summary |>
  filter(term != "(Intercept)") |>
  rename_mlm_terms() |>
  mutate(
    Estimate    = estimate,
    lower_ci    = estimate - 1.96        * std.error,
    upper_ci    = estimate + 1.96        * std.error,
    lower_ci_90 = estimate - qnorm(0.95) * std.error,
    upper_ci_90 = estimate + qnorm(0.95) * std.error,
    Village = recode(Village,
                     "Ampandrana"   = "A",
                     "Mandena"      = "M",
                     "Sarahandrano" = "S")
  )

ggarrange(
  mlm_coef_plot(mlm_full_village, people_levels_full,
                title = "Human Attributes", color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  mlm_coef_plot(mlm_full_village, animal_levels_full,
                title = "Animal Attributes", color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  labels = c("a", "b"), common.legend = TRUE
)
```

## 10d. High-Risk Network - Pooled

```{r plot-hr-pooled, fig.width=10, fig.height=6}
meta_mlm_hr_r <- meta_mlm_hr |> rename_mlm_terms()

ggarrange(
  mlm_coef_plot(meta_mlm_hr_r, people_levels_hr, title = "Human Attributes"),
  mlm_coef_plot(meta_mlm_hr_r, animal_levels_hr, title = "Animal Attributes"),
  labels = c("a", "b"), common.legend = TRUE
)
```

## 10e. High-Risk Network - By Village

```{r plot-hr-village, fig.width=10, fig.height=6}
mlm_hr_village <- mlm_hr_summary |>
  filter(term != "(Intercept)") |>
  rename_mlm_terms() |>
  mutate(
    Estimate    = estimate,
    lower_ci    = estimate - 1.96        * std.error,
    upper_ci    = estimate + 1.96        * std.error,
    lower_ci_90 = estimate - qnorm(0.95) * std.error,
    upper_ci_90 = estimate + qnorm(0.95) * std.error,
    Village = recode(Village,
                     "Ampandrana"   = "A",
                     "Mandena"      = "M",
                     "Sarahandrano" = "S")
  )

ggarrange(
  mlm_coef_plot(mlm_hr_village, people_levels_hr,
                title = "Human Attributes", color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  mlm_coef_plot(mlm_hr_village, animal_levels_hr,
                title = "Animal Attributes", color_var = "Village") +
    scale_color_viridis_d(option = "D", end = 0.8),
  labels = c("a", "b"), common.legend = TRUE
)
```

---

# 11. Coefficient Tables

## 11a. Full Network - Pooled Estimates

```{r table-full}
meta_mlm_full |>
  select(term, Estimate, SE, Lower, Upper, p) |>
  mutate(across(where(is.numeric), ~ round(., 3))) |>
  kable(
    caption   = "Full Network - Pooled MLM Estimates (Fixed-Effects Meta-Analysis)",
    col.names = c("Term", "Estimate", "SE", "Lower 95% CI", "Upper 95% CI", "p-value"),
    row.names = FALSE
  )
```

## 11b. High-Risk Network - Pooled Estimates

```{r table-hr}
meta_mlm_hr |>
  select(term, Estimate, SE, Lower, Upper, p) |>
  mutate(across(where(is.numeric), ~ round(., 3))) |>
  kable(
    caption   = "High-Risk Network - Pooled MLM Estimates (Fixed-Effects Meta-Analysis)",
    col.names = c("Term", "Estimate", "SE", "Lower 95% CI", "Upper 95% CI", "p-value"),
    row.names = FALSE
  )
```

## 11c. By-Village Estimates - Full Network

```{r table-full-village}
mlm_full_summary |>
  filter(term != "(Intercept)") |>
  select(Village, term, estimate, std.error, p.value) |>
  mutate(across(where(is.numeric), ~ round(., 3))) |>
  arrange(term, Village) |>
  kable(
    caption   = "Full Network - Village-Level MLM Estimates",
    col.names = c("Village", "Term", "Estimate", "SE", "p-value"),
    row.names = FALSE
  )
```

## 11d. By-Village Estimates - High-Risk Network

```{r table-hr-village}
mlm_hr_summary |>
  filter(term != "(Intercept)") |>
  select(Village, term, estimate, std.error, p.value) |>
  mutate(across(where(is.numeric), ~ round(., 3))) |>
  arrange(term, Village) |>
  kable(
    caption   = "High-Risk Network - Village-Level MLM Estimates",
    col.names = c("Village", "Term", "Estimate", "SE", "p-value"),
    row.names = FALSE
  )
```

---
# Looking closer at vanilla growers
```{r}
library(tableone)  

# ------------------------------------------------------------------
# Compare vanilla growers vs non-growers on all SES/demographic vars
# One row per participant (not per dyad)
# ------------------------------------------------------------------
vanilla_comparison <- merged_df |>
  distinct(social_netid, .keep_all = TRUE) |>
  mutate(grew_vanilla = factor(grew_vanilla, levels = c(0, 1),
                                labels = c("Did not grow", "Grew vanilla")))

# --- Continuous variables: t-tests ---
continuous_vars <- c("age", "landsize_in_daba", "household_size",
                     "commercial_goods", "house_sol", "school_level_numbered")

ttest_results <- lapply(continuous_vars, function(v) {
  tt <- t.test(reformulate("grew_vanilla", response = v),
               data = vanilla_comparison)
  data.frame(
    Variable  = v,
    Mean_NoVanilla = tt$estimate[1],
    Mean_Vanilla   = tt$estimate[2],
    t         = tt$statistic,
    p         = tt$p.value,
    CI_lower  = tt$conf.int[1],
    CI_upper  = tt$conf.int[2]
  )
}) |> bind_rows()

print(ttest_results)

# --- Categorical variables: chi-squared tests ---
categorical_vars <- c("gender", "school_level", "employment_category", "village")

chisq_results <- lapply(categorical_vars, function(v) {
  tbl <- table(vanilla_comparison[[v]], vanilla_comparison$grew_vanilla)
  cs  <- chisq.test(tbl)
  data.frame(
    Variable = v,
    chi_sq   = cs$statistic,
    df       = cs$parameter,
    p        = cs$p.value
  )
}) |> bind_rows()

print(chisq_results)

# --- Proportions table for categorical vars ---
lapply(categorical_vars, function(v) {
  cat("\n---", v, "---\n")
  print(prop.table(table(vanilla_comparison[[v]],
                         vanilla_comparison$grew_vanilla), margin = 2))
})

# --- Animal contact: do non-growers contact different animals? ---
# Total edge count per person by vanilla status
flat_full_ergm |>
  group_by(from) |>
  summarise(
    total_contacts = sum(edge == 1),
    grew_vanilla   = first(grew_vanilla),
    village        = first(village),
    .groups = "drop"
  ) |>
  mutate(grew_vanilla = factor(grew_vanilla, levels = c(0, 1),
                                labels = c("Did not grow", "Grew vanilla"))) |>
  group_by(grew_vanilla) |>
  summarise(
    n             = n(),
    mean_contacts = mean(total_contacts),
    sd_contacts   = sd(total_contacts),
    .groups = "drop"
  ) |>
  print()

# Contact rates by specific animal category
flat_full_ergm |>
  filter(!is.na(animal_category)) |>
  mutate(grew_vanilla = factor(grew_vanilla, levels = c(0, 1),
                                labels = c("Did not grow", "Grew vanilla"))) |>
  group_by(grew_vanilla, animal_category) |>
  summarise(contact_rate = mean(edge == 1), .groups = "drop") |>
  pivot_wider(names_from = grew_vanilla, values_from = contact_rate) |>
  print()
```
