---
title: "Self-Reported Network Descriptive Stats"
author: "Lev Kolinski"
date: "`r Sys.Date()`"
output: html_document
---

## Set Up

```{r setup}
library(here)
library(ineq)
library(kableExtra)
library(tidyverse)

source(here("self_reported_networks/The_ReUp/1. Animal Interaction Bipartite Full Network.R"))
source(here("self_reported_networks/The_ReUp/3. Animal Interaction Bipartite High Risk.R"))


# LOADING DFs----
library(tidyverse)
demographic_health_data <- read.csv("/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Demographic_Health.csv")
animal_interaction_data <- read.csv("/Users/levkolinski/Library/CloudStorage/Box-Box/EEID_Data_public/clean_data_tables/NIH Human Surveys/Survey_Animal_Interaction.csv")


# cleaning data, selecting only columns that I want, converting binaries to 0 and 1
demographic_health_data_cleaned <- demographic_health_data |> 
  mutate(across(starts_with(c("own_", "banking", "grew_crops",
                              "crops")),
                ~ifelse(. == "Yes", 1L, 0L))) |> 
  select(survey_date, social_netid, village,age,gender,school_level, household_size, main_activity, 
         starts_with("material_"), starts_with("own_"),crops_vanilla, landsize_in_daba)


animal_interaction_data_cleaned <- animal_interaction_data |>  
  mutate(across(-starts_with("social_netid"), ~ifelse(. == "Yes", 1L, 0L))) |> 
  select(-starts_with("symptoms"),-starts_with("action_"), -contains("frequency"))


merged_df <- inner_join(demographic_health_data_cleaned, animal_interaction_data_cleaned, by = "social_netid")

merged_df <- suppressWarnings(merged_df |> 
                                        mutate(material_wall_index = 0) %>%
                                        mutate(material_wall_index = if_else(material_wall == "bamboo", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "rafia", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "ravenala", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "mud", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "compacted_earth", 0, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "wood_planks", 1, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "brick_unfired", 2, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "brick_fired", 2, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "metal_sheets", 3, material_wall_index)) %>%
                                        mutate(material_wall_index = if_else(material_wall == "cement", 4, material_wall_index)) %>%
                                        mutate(material_wall_index = scale(material_wall_index)) |> 
                                        
                                        # roof construction
                                        mutate(material_roof_index = 0) %>%
                                        mutate(material_roof_index = if_else(material_roof == "bamboo", 0, material_roof_index)) %>%
                                        mutate(material_roof_index = if_else(material_roof == "thatch", 0, material_roof_index)) %>%
                                        mutate(material_roof_index = if_else(material_roof == "metal_sheets", 1, material_roof_index)) %>%
                                        mutate(material_roof_index = if_else(material_roof == "cement", 2, material_roof_index)) %>%
                                        mutate(material_roof_index = scale(material_roof_index)) |> 
                                        
                                        # floor construction
                                        mutate(material_floor_index = 0) %>%
                                        mutate(material_floor_index = if_else(material_floor == "dirt", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "bamboo", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "rafia", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "ravinala", 0, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "wood_planks", 1, material_floor_index)) %>%
                                        mutate(material_floor_index = if_else(material_floor == "cement", 2, material_floor_index)) %>%
                                        mutate(material_floor_index = scale(material_floor_index)) |> 
                                        
                                        # house index
                                        mutate(house_sol = material_wall_index + material_roof_index + material_floor_index) |> 
                                        mutate(house_sol = scale(house_sol)) |> 
                                        
                                        # commercial goods
                                        mutate(commercial_goods = own_cellphone + own_tv + own_bicycle + own_refrigerator +
                                                 own_motorcycle + own_computer + own_generator) |> 
                                        mutate(commercial_goods = scale(commercial_goods)) |> 
                                        
                                        # other things
                                        mutate(employment_category = if_else(grepl("farm_crops", main_activity) | grepl("farm_mixed", main_activity), "Farmer", 
                                                                             ifelse(grepl("student", main_activity) | grepl("teacher", main_activity), "Education",
                                                                                    ifelse(grepl("unemployed", main_activity) | grepl("retired_teacher", main_activity), "Unemployed",
                                                                                           "Other")))) |> 
                                        mutate(school_level_numbered = ifelse(school_level == "None", 0L, 
                                                                              ifelse(school_level == "Primary", 1L,
                                                                                     ifelse(school_level == "Secondary", 2L,
                                                                                            ifelse(school_level == "Higher", 3L,
                                                                                                   NA))))) |> 
                                        rename(grew_vanilla = crops_vanilla) |> 
                                        mutate(age = scale(age)) |> 
                                        mutate(landsize_in_daba = scale(landsize_in_daba)) |> 
                                        mutate(household_size = scale(household_size)) |> 
                                        mutate(school_level_numbered = scale(school_level_numbered))) |> 
  drop_na(house_sol, commercial_goods, grew_vanilla, age, landsize_in_daba, household_size, school_level_numbered, gender)
                                
merged_df$survey_date <- as.Date(merged_df$survey_date)


n_respondents<-length(unique(merged_df$social_netid))
n_respondents

```

## FULL NETWORK

### Network Descriptive Stats Table 
```{r}

# Function to calculate degree statistics and network properties for each village
calculate_network_stats <- function(graph, village_name) {
  human_degrees <- V(graph)$degree[V(graph)$type]  # Degrees for humans
  animal_degrees <- V(graph)$degree[!V(graph)$type]  # Degrees for animals
  
  
  # Collect statistics
  stats <- data.frame(
    Village = village_name,
    Number_Human_Nodes = sum(V(graph)$type == TRUE),
    Mean_Human_Degree = mean(human_degrees),
    SD_Human_Degree = sd(human_degrees),
    Min_Human_Degree = min(human_degrees),
    Max_Human_Degree = max(human_degrees),
    Number_Animal_Nodes = sum(V(graph)$type == FALSE),
    Mean_Animal_Degree = mean(animal_degrees),
    SD_Animal_Degree = sd(animal_degrees),
    Min_Animal_Degree = min(animal_degrees),
    Max_Animal_Degree = max(animal_degrees),
    Num_Edges = ecount(graph),
    Density = igraph::edge_density(graph)
  )
  
  return(stats)
}

# List of graph objects and corresponding village names
graphs <- list(
  "Ampandrana Andatsakala" = bipartite_graph_full_network_ampandrana_andatsakala,
  "Mandena" = bipartite_graph_full_network_mandena,
  "Sarahandrano" = bipartite_graph_full_network_sarahandrano
)

# Iterate over graphs and village names to calculate stats
village_stats_list <- lapply(names(graphs), function(village_name) {
  calculate_network_stats(graphs[[village_name]], village_name)
})

# Combine the results into a single data frame
village_stats <- bind_rows(village_stats_list)
village_stats$Village <- factor(village_stats$Village, levels = c("Mandena","Sarahandrano", "Ampandrana Andatsakala"))
village_stats <- village_stats[order(village_stats$Village), ]


# Print the results as a table using knitr::kable for clean display in R Markdown
kable(village_stats, 
      caption = "Full Network Statistics Comparison Across Villages",
      col.names = c("Village","Number Human Nodes", "Mean Human Degree", "SD Human Degree", "Min Human Degree", 
                    "Max Human Degree", "Number Animal Nodes", "Mean Animal Degree", "SD Animal Degree", "Min Animal Degree",
                    "Max Animal Degree",
                    "Number of Edges", "Density"),
      digits = 4,
      row.names = FALSE) 
```

```{r}

# FULL NETWORKS ----

### get all unique animals
animals <- c("rodents", "poultry", "dogs", "cats", "domestic_pigs", "goats_sheep",
             "cows", "wild_birds", "bush_pigs", "lemurs", "tenrecs", "carnivores")

# Get all unique humans from merged survey df
humans <- unique(merged_df$social_netid)

# Create all possible combinations of humans and animals
edge_df <- expand.grid(from = humans, to = animals, stringsAsFactors = FALSE)
edge_df<- edge_df |> 
  arrange(from)

# Initialize edge column to 0
edge_df$edge <- 0

# Loop over each animal type
for (animal in animals) {
  
  animal_cols <- colnames(merged_df)[grepl(animal, colnames(merged_df))]
  
  if (length(animal_cols) > 0) {
    # Check if any contact occurred for each person
    merged_df[[paste0("has_", animal, "_contact")]] <- rowSums(merged_df[animal_cols], na.rm = TRUE) > 0
    
    # Update the edge_df accordingly
    edge_df$edge[edge_df$to == animal & 
                   edge_df$from %in% merged_df$social_netid[merged_df[[paste0("has_", animal, "_contact")]]]] <- 1
  }
}

# adding in animal_category
domesticated_animals <- c("poultry", "dogs", "domestic_pigs", "cows", "goats_sheep", "cats")
rodent_animals <- "rodents"
wild_animals <- c("bush_pigs", "carnivores", "tenrecs", "wild_birds", "lemurs")

edge_df<-edge_df |> 
  mutate(animal_category = ifelse(to %in% domesticated_animals, "domesticated",
                                  ifelse(to %in% rodent_animals, "rodent",
                                         ifelse(to%in%wild_animals, "wild", NA))))

# bring in human survey data
bipartite_graph_full_network_meta_data<-igraph::as_data_frame(bipartite_graph_full_network ,
                                                              what = "vertices") |> 
  rename(from = name)

# merge with edge list
full_network_df<-merge(edge_df,bipartite_graph_full_network_meta_data,
                       by = "from")
```

### Only X% of Individuals Contribute 50% of All INTERACTIONS
```{r}
full_network_df |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  arrange(desc(total_edges)) |> 
  mutate(
    cum_people = row_number() / n(),
    cum_edges  = cumsum(total_edges) / sum(total_edges)
  ) |> 
  filter(cum_edges >= 0.50) |> 
  slice(1) |> 
  summarise(
    percent_of_people_for_half_interactions = 100 * cum_people
  )
```

### Average number interactions per person
```{r}
full_network_df |> 
  filter(!is.na(edge)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    .groups = "drop"
  ) |> 
  summarise(
    average_interactions_per_person = mean(total_edges),
    sd_interactions = sd(total_edges),
    n_people = n()
  )
```

### Land Size Differences
```{r}
landsize_quintile_edges <- full_network_df |> 
  filter(!is.na(edge), !is.na(land_size)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    land_size = first(land_size),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(land_size),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = landsize_quintile_edges)
```

### COMMERCIAL GOODS DIFFERENCES 
```{r}
commercial_goods_quintile_edges <- full_network_df |> 
  filter(!is.na(edge), !is.na(commercial_goods)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    commercial_goods = first(commercial_goods),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(commercial_goods),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = commercial_goods_quintile_edges)
```


### HOUSE MATERIALS DIFFERENCES 
```{r}
house_mat_quintile_edges <- full_network_df |> 
  filter(!is.na(edge), !is.na(house_sol)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    house_sol = first(house_sol),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(house_sol),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = house_mat_quintile_edges)
```

### VANILLA DIFFERNCES
```{r}
edge_summary_vanilla <- full_network_df |> 
  filter(!is.na(edge), !is.na(grew_vanilla)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    grew_vanilla = first(grew_vanilla),
    .groups = "drop"
  )

t.test(total_edges ~ grew_vanilla, data = edge_summary_vanilla)
```

## High-Risk Network

### Network Descriptive Stats Table 
```{r}
graphs_high_risk <- list(
  "Ampandrana Andatsakala" = bipartite_graph_high_risk_ampandrana_andatsakala,
  "Mandena" = bipartite_graph_high_risk_mandena,
  "Sarahandrano" = bipartite_graph_high_risk_sarahandrano
)

# Iterate over graphs_high_risk and village names to calculate stats
village_stats_list_high_risk <- lapply(names(graphs_high_risk), function(village_name) {
  calculate_network_stats(graphs_high_risk[[village_name]], village_name)
})

# Combine the results into a single data frame
village_stats_high_risk <- bind_rows(village_stats_list_high_risk)
village_stats_high_risk$Village <- factor(village_stats_high_risk$Village, levels = c("Mandena","Sarahandrano", "Ampandrana Andatsakala"))
village_stats_high_risk <- village_stats_high_risk[order(village_stats_high_risk$Village), ]

# Print the results as a table using knitr::kable for clean display in R Markdown
kable(village_stats_high_risk, 
      caption = "Full Network Statistics Comparison Across Villages",
      col.names = c("Village", "Number Human Nodes", "Mean Human Degree", "SD Human Degree", "Min Human Degree", 
                    "Max Human Degree", "Number Animal Nodes", "Mean Animal Degree", "SD Animal Degree", "Min Animal Degree",
                    "Max Animal Degree",
                    "Number of Edges", "Density"),
      digits = 4,
      row.names = FALSE)

```


```{r}

# get all unique animals
animals <- c("rodents", "poultry", "dogs", "cats", "domestic_pigs", "goats_sheep",
             "cows", "wild_birds", "bush_pigs", "lemurs", "tenrecs", "carnivores")

# Get all unique humans from merged survey df
humans <- unique(merged_df$social_netid)

# Create all possible combinations of humans and animals
edge_df_high_risk <- expand.grid(from = humans, to = animals, stringsAsFactors = FALSE)
edge_df_high_risk<- edge_df_high_risk |> 
  arrange(from)

# Initialize edge column to 0
edge_df_high_risk$edge <- 0

# Loop over each animal type
for (animal in animals) {
  
  animal_cols <- colnames(merged_df)[grepl(animal, colnames(merged_df))&grepl("feces|shared_water|scratched|cooked|raw|sick|slaughtered", 
                                                                              colnames(merged_df))]
  
  if (length(animal_cols) > 0) {
    # Check if any contact occurred for each person
    merged_df[[paste0("has_", animal, "_contact")]] <- rowSums(merged_df[animal_cols], na.rm = TRUE) > 0
    
    # Update the edge_df accordingly
    edge_df_high_risk$edge[edge_df_high_risk$to == animal & 
                             edge_df_high_risk$from %in% merged_df$social_netid[merged_df[[paste0("has_", animal, "_contact")]]]] <- 1
  }
}



# adding in animal_category
domesticated_animals_1 <- c("poultry", "domestic_pigs", "cows")
domesticated_animals_2 <- c( "dogs","goats_sheep", "cats")
rodent_animals <- "rodents"
wild_animals <- c("bush_pigs", "carnivores", "tenrecs", "wild_birds", "lemurs")

edge_df_high_risk<-edge_df_high_risk |> 
  mutate(animal_category = ifelse(to %in% domesticated_animals_1, "domesticated_1",
                                  ifelse(to %in% domesticated_animals_2, "domesticated_2",
                                         ifelse(to%in%rodent_animals, "rodent", 
                                                ifelse(to%in%wild_animals, "wild", NA)))))

# bring in human survey data
bipartite_graph_high_risk_network_meta_data<-igraph::as_data_frame(bipartite_graph_high_risk ,
                                                                   what = "vertices") |> 
  rename(from = name)

# merge with edge list
high_risk_network_df<-merge(edge_df_high_risk,bipartite_graph_high_risk_network_meta_data,
                            by = "from")

```


### Only X% of Individuals Contribute 50% of All INTERACTIONS
```{r}
high_risk_network_df |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  arrange(desc(total_edges)) |> 
  mutate(
    cum_people = row_number() / n(),
    cum_edges  = cumsum(total_edges) / sum(total_edges)
  ) |> 
  filter(cum_edges >= 0.50) |> 
  slice(1) |> 
  summarise(
    percent_of_people_for_half_interactions = 100 * cum_people
  )

```

### Average number interactions per person
```{r}
high_risk_network_df |> 
  filter(!is.na(edge)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    .groups = "drop"
  ) |> 
  summarise(
    average_interactions_per_person = mean(total_edges),
    sd_interactions = sd(total_edges),
    n_people = n()
  )

```


### LAND SIZE DIFFERENCES
```{r}
landsize_quintile_edges <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(land_size)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    land_size = first(land_size),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(land_size),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = landsize_quintile_edges)
```

### COMMERCIAL GOODS DIFFERENCES
```{r}
commercial_goods_quintile_edges <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(commercial_goods)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    commercial_goods = first(commercial_goods),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(commercial_goods),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = commercial_goods_quintile_edges)
```

### HOUSE MATERIALS DIFFERENCES 
```{r}
house_mat_quintile_edges <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(house_sol)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    house_sol = first(house_sol),
    .groups = "drop"
  ) |> 
  mutate(rank = percent_rank(house_sol),
         quintile = case_when(
           rank <= 0.20 ~ "bottom_20",
           rank >= 0.80 ~ "top_20",
           TRUE ~ NA
         )) |> 
  filter(!is.na(quintile))

t.test(total_edges ~ quintile, data = house_mat_quintile_edges)

```

### VANILLA DIFFERNCES
```{r}
edge_summary_vanilla <- high_risk_network_df |> 
  filter(!is.na(edge), !is.na(grew_vanilla)) |> 
  group_by(from) |> 
  summarise(
    total_edges = sum(edge == 1),
    grew_vanilla = first(grew_vanilla),
    .groups = "drop"
  )

t.test(total_edges ~ grew_vanilla, data = edge_summary_vanilla)
```


